---
title: "State wide NO~3~^-^ leaching results"
author: "Elke Brandes"
date: "November 15, 2016"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## DNDC - Model assumptions

* Iowa, subfield resolution

* 4 model runs:

    1. Status quo (corn, soybean), 2012-2015
    
    2. Switchgrass, low yield (7,500 kg/ha), 2006-2015
    
    3. Switchgrass, med yield (10,000 kg/ha), 2006-2015
    
    4. Switchgrass, high yield (12,500 kg/ha), 2006-2015
    
## DNDC - Model assumptions cont'd

- Row crop fertilization rates are constant, varying by year (optimal return on N)
- Switchgrass: 50 kg N/ha
- Switchgrass C/N above ground biomass = 75
- Switchgrass C/N below ground biomass = 60

## Results - Outline

1. Row crop profitability
2. NO~3~^-^ leaching in Iowa
3. Switchgrass integration scenarios
4. Spatial distribution of NO~3~^-^ leaching 



```{r intro, message=FALSE, warning=FALSE}
library('RPostgreSQL') # PostgreSQL database connection
library("ggplot2") # for the figures
library("maps") # to plot maps, contains country/state/county outlines
library("gridGraphics") # needed for the function unit in ggplot
library("tidyverse") # to make tidy tables, includes dplyr needed for PostgreSQL database connection
library('knitr') # to use kable() function

library(rgdal) # to read in shape files
library(ggmap)
library(scales)
library(RColorBrewer)
library("sp") # to use merge with a spatial data frame

library("tmap")

pw <- xxxx
isuag_db <- src_postgres(dbname = 'isuag',
                           host = 'isu-ag-db.agsolver.com',
                           port = 5432,
                           user = 'isuag',
                           password = pw)
pw <- 
drv <- dbDriver("PostgreSQL") # loads the PostgreSQL driver
con <- dbConnect(drv, dbname = "isuag",  # creates a connection to the postgres database
                 host = "isu-ag-db.agsolver.com", port = 5432,
                 user = "isuag", password = pw)
# note that "con" will be used later in each connection to the database

cgsb_leach <-tbl(isuag_db, "06_cgsb_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
swg_low <-tbl(isuag_db, "06_swg_7500_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
swg_med <-tbl(isuag_db, "06_swg_10000_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
swg_hi <-tbl(isuag_db, "06_swg_12500_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()

data_list <- list(cgsb_leach, swg_low, swg_med, swg_hi)

theme_b_border <- function (base_size = 12, base_family = "") 
{
  theme_grey(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(size = rel(0.8), margin = margin(r=10)), #margin 
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(0.15, "cm"),
      legend.key = element_rect(colour = NA), panel.background = element_rect(fill = "white", 
      colour = NA), panel.border = element_rect(fill = NA, 
      colour = "black"), panel.grid.major = element_line(colour = NA, 
      size = 0.2), panel.grid.minor = element_line(colour = NA, 
      size = 0.5), strip.background = element_rect(fill = "grey80", 
      colour = "grey50", size = 0.2))
}

scenario <- c(rep("cgsb", sum(sapply(data_list[1],nrow))), rep("swg_low", sum(sapply(data_list[2],nrow))), rep("swg_med", sum(sapply(data_list[3],nrow))), rep("swg_hi", sum(sapply(data_list[4],nrow)))) %>%
factor(levels= c('cgsb', 'swg_low', 'swg_med', 'swg_hi')) # make it a factor so that it is plotted in the right order in the facets

leaching <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   leaching <- append(leaching, data_list[[i]][,1])
  
area <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   area <- append(area, data_list[[i]][,2])

leach <- data.frame(scenario, leaching, area = area*1e-6) # this factor to show Mha

```

# 1. Row crop profitability

## Mean profit distribution

```{r profit_distribution, fig.cap="Figure 1: Distribution of mean profitability 2012-2015."}
profit = tbl(isuag_db, "01_profit_mean_2012_2015_aggregated") %>%
  as.data.frame()
profit$sum_ha <- profit$sum_ha * 1e-6  # convert to Mha

ggplot(profit, aes(profit_mean_ha_rounded, weight = sum_ha)) + 
  geom_histogram(binwidth=20, alpha = .5, position="identity") +
  theme_b_border() +
  labs(x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1})),
       y=expression(Area~(Mha))) +
  scale_x_continuous() +
  scale_y_continuous() +
  geom_vline(aes(xintercept= -150, linetype= "dashed"),  show.legend = FALSE) 
```

## Spatial distribution of highly unprofitable areas

```{r county_shapefile, include=FALSE}
counties <- readOGR(dsn="Z:/ElkeBrandes/projects/01 subfield profit/data analysis/02 GIS/government_units/iowa_cnty/IA_cnty/county.shp",layer="county")

```


```{r profit_spatial, fig.width=8.5, fig.height=3.5, fig.cap="Figure 2: Areas below a mean 2012-2015 profitability of -150 US$ ha^-1^ per county."}
names(counties)[9] <- "fips" 
profit_county <- tbl(isuag_db, "07_swg_area_counties_po") %>%
  as.data.frame() 

 for (i in 1:length(profit_county$fips)) {
   profit_county$fips[i] <- sub("IA", "19", profit_county$fips[i])
    }

# profit_county$fips <- as.factor(profit_county$fips)
counties@data <- left_join(counties@data, profit_county, by = "fips")

counties_df <- fortify(counties)
counties$id <- row.names(counties) 
counties_profit_df <- left_join(counties_df, counties@data)

# add a spatial data frame with the county centroids, to be used to plot the total leaching N per county:
centroids <- getSpPPolygonsLabptSlots(counties) %>%
  as.data.frame() 

centroids$id <- as.character(0:98)

# join the total N leaching data to the centroids 
counties_profit_df <- left_join(counties_profit_df, centroids)

# create a theme without axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

ggplot() +
  geom_polygon(data = counties_profit_df, aes(long, lat, group = group, fill=area_in_swg_perc), color = "white") +
  geom_point(data = counties_profit_df, aes(V1, V2, size = area_in_swg), color = "orange") +
  labs(size = expression(Total~area~below~-150~"US$"~ha^{-1}~(ha))) +
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(Relative~area~below~-150~"US$"~ha^{-1}~("%"))) +
  scale_fill_gradient(low = "white", high="dark blue", guide = "colourbar") +
  ggtitle("Highly unprofitable area per county")

```

# 2. NO~3~^-^ leaching in Iowa

## Distributions of NO~3~^-^ leaching
```{r histograms, warning=FALSE, message=FALSE, cache=TRUE, fig.cap="Figure 3: Distributions of NO~3~^-^ leaching"}
scenario <- c(rep("cgsb", sum(sapply(data_list[1],nrow))), rep("swg_low", sum(sapply(data_list[2],nrow))), rep("swg_med", sum(sapply(data_list[3],nrow))), rep("swg_hi", sum(sapply(data_list[4],nrow)))) %>%
factor(levels= c('cgsb', 'swg_low', 'swg_med', 'swg_hi')) # make it a factor so that it is plotted in the right order in the facets

leaching <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   leaching <- append(leaching, data_list[[i]][,1])
  
area <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   area <- append(area, data_list[[i]][,2])

   
leach <- data.frame(scenario, leaching, area = area*1e-6) # this factor to show Mha

scenarios <- c('cgsb'= "Status Quo (corn/soybeans)", 'swg_low' = "Low yielding switchgrass", 'swg_med' = "Medium yielding switchgrass", 'swg_hi' = "High yielding switchgrass")

ggplot(leach, aes(leaching, weight = area)) + 
  geom_histogram(binwidth=2, alpha = .5, position="identity") +
  theme_b_border() +
  labs(x=expression(Mean~NO[3]^{"-"}~leaching~2006-2015~(kg~N~ha^{-1})),
       y=expression(Area~(Mha))) +
  scale_x_continuous() +
  scale_y_continuous() +
  theme(legend.position="bottom") +
  facet_wrap( ~ scenario, ncol=2, labeller = as_labeller(scenarios)) 
```

 

## Correlation between profitability and NO~3~^-^ leaching
```{sql, connection=con, cache=TRUE, include=FALSE}
SELECT setseed(0.5)
```
```{sql, connection=con, message=FALSE, cache=TRUE}
DROP TABLE IF EXISTS "r_dndc_clumu_cgsb_swg_sample";
```
```{sql, connection=con, message=FALSE, cache=TRUE}
CREATE TABLE "r_dndc_clumu_cgsb_swg_sample" AS SELECT mean_profit_ha, ave_no3_leach_ha_cgsb FROM "05_dndc_clumu_cgsb_swg" ORDER BY random() LIMIT 10000
```

```{r data_sample}
sample = tbl(isuag_db, "r_dndc_clumu_cgsb_swg_sample") %>%
  as.data.frame(sample)
```

```{r scatterplot, cache=TRUE, fig.show="hold", fig.width=4, fig.height=4, fig.cap="Figure 4: Scatterplot of a subsample (n=10,000) of profitability and corn/soy NO~3~^-^ leaching data on subfield areas in Iowa."}
ggplot() +
  geom_point(data = sample, aes(x = mean_profit_ha, y = ave_no3_leach_ha_cgsb)) +
  theme_b_border() +
  labs(y=expression(Mean~NO[3]^{"-"}~leaching~2006-2015~(kg~N~ha^{-1})),
       x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1}))) +
  geom_vline(aes(xintercept= -150, linetype= "dashed", color = "red"),  show.legend = FALSE) +
  geom_hline(aes(yintercept= 60, linetype= "dotted", color = "red"),  show.legend = FALSE)

ggplot() +
  geom_point(data = sample, aes(x = mean_profit_ha, y = ave_no3_leach_ha_cgsb)) +
  theme_b_border() +
  labs(y=expression(Mean~NO[3]^{"-"}~leaching~2006-2015~(kg~N~ha^{-1})),
       x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1}))) +
  scale_x_continuous() +
  scale_y_continuous(limits=c(0,20)) +
  geom_vline(aes(xintercept= -150, linetype= "dashed", color = "red"),  show.legend = FALSE)
```


# 3. Switchgrass integration scenarios

## Varying the thresholds

```{r swg_areas_po, cache=TRUE}
swg_areas_po <- vector(mode="numeric", length=0)
# profit <- tbl(isuag_db, "05_dndc_clumu_cgsb_swg")
profit_cutoffs <- c(-500, -400, -300, -200, -150, -100, -50, 0, 50, 100, 150)
 for (i in 1:length(profit_cutoffs)) {
    query <- paste("SELECT SUM(clumuha) FROM \"05_dndc_clumu_cgsb_swg\" WHERE mean_profit_ha < ", as.character(profit_cutoffs[i]))
    sum_area <- tbl(isuag_db, sql(query)) %>%
      as.data.frame()
    sum_area <- sum_area[1,1]
    swg_areas_po <- append(swg_areas_po, sum_area) 
  }

sum_leaching <- tbl(isuag_db, "05_dndc_no3_leach_sums_iowa_dispr_benefits") %>%
  as.data.frame() %>%
  gather(corn_soybeans, swg_7500_1, swg_7500_2, swg_7500_3, swg_7500_4, swg_7500_5, swg_7500_6, swg_7500_7, swg_7500_8, swg_7500_9, swg_7500_10, swg_7500_11, swg_10000_1, swg_10000_2, swg_10000_3, swg_10000_4, swg_10000_5, swg_10000_6, swg_10000_7, swg_10000_8, swg_10000_9, swg_10000_10, swg_10000_11, swg_12500_1, swg_12500_2, swg_12500_3, swg_12500_4, swg_12500_5, swg_12500_6, swg_12500_7, swg_12500_8, swg_12500_9, swg_12500_10, swg_12500_11, key = swg_scenarios, value=ave_no2_leach)

```

```{r swg_areas_wo, cache=TRUE}
swg_areas_wo <- vector(mode="numeric", length=0)
leach_cutoff <- c(100, 90, 80, 70, 60, 50)
 for (i in 1:length(leach_cutoff)) {
    query <- paste("SELECT SUM(clumuha) FROM \"05_dndc_clumu_cgsb_swg\" WHERE ave_no3_leach_ha_cgsb > ", as.character(leach_cutoff[i]))
    sum_area <- tbl(isuag_db, sql(query)) %>%
      as.data.frame()
    sum_area <- sum_area[1,1]
    swg_areas_wo <- append(swg_areas_wo, sum_area) 
  }
swg_yield <- as.factor(c("NA",rep("7500",11), rep("10000",11),rep("12500",11)))
profit_cutoff <- c(NA, rep(profit_cutoffs,3))
swg_area_po <- c(0,rep(swg_areas_po,3))
swg_area_po_percent <- (swg_area_po/9374362)*100
N_leaching <- sum_leaching[,2]
N_reduction <- abs((N_leaching - N_leaching[1])/N_leaching[1])*100

relation_po <- data.frame(swg_yield, profit_cutoff, swg_area_po_percent, N_reduction)

swg_area_wo_percent <- (swg_areas_wo/9374362)*100

relation_wo <- data.frame(leach_cutoff, swg_area_wo_percent)
```

```{r thresholds, fig.show="hold", fig.height= 3, fig.width=3.5, fig.cap="Figure 5: Relation between the set thresholds in the profitability (left panel) and water quality optimized (right panel) scenario and the resulting area in switchgrass."}
ggplot() +
  geom_point(data = relation_po, aes(x = profit_cutoff, y = swg_area_po_percent)) +
  theme_b_border() +
  labs(x=expression(Profitability~threshold~("US$"~ha^{-1})),
       y=expression(Area~"in"~switchgrass~("%"))) +
  scale_x_continuous(breaks=seq(from=-500, to=200, by= 100)) +
  scale_y_continuous(limits= c(0,65)) +
    geom_vline(aes(xintercept= -150, linetype= "dotted", color = "red"),  show.legend = FALSE)

ggplot() +
  geom_point(data = relation_wo, aes(x = leach_cutoff, y = swg_area_wo_percent)) +
  theme_b_border() +
    labs(x=expression(NO[3]^{"-"}~leaching~threshold~(kg~ha^{-1})),
       y=expression(Area~"in"~switchgrass~("%"))) +
  scale_x_reverse(breaks=seq(from=50, to=100, by= 10)) +
  scale_y_continuous(limits= c(0,65)) +
    geom_vline(aes(xintercept= 60, linetype= "dotted", color = "red"),  show.legend = FALSE)
```

## State-wide NO~3~^-^ leaching

```{r total_leaching, cache=TRUE}
leach_tot <-tbl(isuag_db, "05_dndc_no3_leach_sums_iowa_scenarios")  %>%
  as.data.frame()
names(leach_tot) <- c("Status Quo", "PO 7500", "PO 10000", "PO 12500","WO 7500", "WO 10000", "WO 12500")

leach_red <- data.frame(status_quo = "NA",
                        po_7500 = (leach_tot[1,1]-leach_tot[1,2])*100/leach_tot[1,1],
                        po_10000 = (leach_tot[1,1]-leach_tot[1,3])*100/leach_tot[1,1],
                        po_12500 = (leach_tot[1,1]-leach_tot[1,4])*100/leach_tot[1,1],
                        wo_7500 = (leach_tot[1,1]-leach_tot[1,5])*100/leach_tot[1,1],
                        wo_10000 = (leach_tot[1,1]-leach_tot[1,6])*100/leach_tot[1,1],
                        wo_12500 = (leach_tot[1,1]-leach_tot[1,7])*100/leach_tot[1,1])
names(leach_red) <- c("Status Quo", "PO 7500", "PO 10000", "PO 12500","WO 7500", "WO 10000", "WO 12500")

leach_table <- rbind(leach_tot[1,],leach_red[1,])

kable(leach_table, digits = 0, col.names=c("Status Quo","PO 7500", "PO 10000", "PO 12500","WO 7500", "WO 10000", "WO 12500"))
```

- PO scenario: Areas losing 150 US$ ha^-1^ or more are in switchgrass
- WO scenario: Areas leaching 60 kg ha^-1^ or more are in switchgrass


## Disproportionate benefits?

```{r relations, cache=TRUE, fig.show="hold", fig.height= 3, fig.width=3.5, fig.cap="Figure 6: Increase in NO~3~^-^ leaching reduction with increasing area in switchgrass (left panel), and with increasing profitability threshold (right panel)."}

ggplot() +
  geom_point(data = relation_po, aes(x = swg_area_po_percent, y = N_reduction, color=swg_yield)) +
  theme_b_border() +
  labs(x=expression(Area~"in"~switchgrass~("%")),
       y=expression(NO[3]^{"-"}~leaching~reduction~("%"~N))) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete(name="Switchgrass\nYield (kg/ha)",  guide = FALSE)

ggplot() +
  geom_point(data = relation_po, aes(x = profit_cutoff, y = N_reduction, color=swg_yield)) +
  theme_b_border() +
  labs(x=expression(Profitability~threshold~("US$"~ha^{-1})),
       y=expression(NO[3]^{"-"}~leaching~reduction~("%"~N))) +
  scale_x_continuous(breaks=seq(from=-500, to=200, by= 100)) +
  scale_y_continuous() +
  scale_color_discrete(name="Switchgrass\nYield") +
  theme(legend.position=c(0.2,0.6))

```

#4. Spatial distribution of   NO~3~^-^ leaching 

## NO~3~^-^ leaching from row crops per county

```{r counties_attributes, include=FALSE}
 counties_leach = tbl(isuag_db, "07_dndc_counties") %>%
   as.data.frame()
 # change the data in the column named "FIPS", so that is contains the county identifiers in the same format as the spatial file:
 for (i in 1:length(counties_leach$fips)) {
   counties_leach$fips[i] <- sub("IA", "19", counties_leach$fips[i])
    }
 # make it a factor
  counties_leach$fips <- as.factor(counties_leach$fips)

  counties <- readOGR(dsn="Z:/ElkeBrandes/projects/01 subfield profit/data analysis/02 GIS/government_units/iowa_cnty/IA_cnty/county.shp",layer="county")
  names(counties)[9] <- "fips" # change to lower case so that we can join on this field later on
counties@data <- left_join(counties@data, counties_leach, by = "fips")
counties_df <- fortify(counties)
counties$id <- row.names(counties) # assign a column with the variable ID that is equal to the row names (=numbers), sarting with 0.
# join the attributes back to the data frame, to create a dataframe used to map the baseline (corn/soybean):
counties_baseline_df <- left_join(counties_df, counties@data)
  
```


```{r plot_baseline_map, fig.height=3.5 , fig.width=9 , fig.cap="Figure 7: County mean annual NO~3~^-^ leaching, averaged over 2012-2015, from corn and soybean fields in Iowa."} 

# create a theme without axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# add a spatial data frame with the county centroids, to be used to plot the total leaching N per county:
centroids <- getSpPPolygonsLabptSlots(counties) %>%
  as.data.frame() 
centroids$id <- as.character(0:98)

# join the total N leaching data to the centroids 
counties_baseline_df <- left_join(counties_baseline_df, centroids)


ggplot() +
  geom_polygon(data = counties_baseline_df, aes(long, lat, group = group, fill=ave_no3_leach_cgsb), color = "white") +
  geom_point(data = counties_baseline_df, aes(V1, V2, size = tot_ave_no3_leach_cgsb), color = "yellow") +
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(Average~NO[3]^{"-"}~leaching~(kg~N~ha^{-1}))) +
  labs(size = expression(Total~NO[3]^{"-"}~leaching~(Mg~N~ha^{-1}))) +
  scale_fill_gradient(low = "white", high="#009E73", guide = "colourbar")

```

## NO~3~^-^ leaching from row crops integrated with switchgrass per county
```{r plot_leach_maps_yields,fig.width=9, fig.height=3.5, fig.cap="Figure 8: County mean annual NO~3~^-^ leaching, averaged over 2012-2015, from corn and soybean fields integrated with switchgrass."}
# set up the data frame 
counties_swg <- data.frame(id = rep(counties$id,3),
                         scenario = c(rep("swg_7500", length(counties$id)), rep("swg_10000", length(counties$id)), 
                                                                               rep("swg_12500", length(counties$id))),
                         tot_no3_leach = c(counties$tot_ave_no3_leach_7500, counties$tot_ave_no3_leach_10000,
                                           counties$tot_ave_no3_leach_12500),
                         ave_no3_leach = c(counties$ave_no3_leach_7500, counties$ave_no3_leach_10000, counties$ave_no3_leach_12500),
                         leach_red = c(counties$no3_leach_change_7500, counties$no3_leach_change_10000,
                                       counties$no3_leach_change_12500))
counties_swg$scenario <- factor(counties_swg$scenario, levels = c("swg_7500", "swg_10000", "swg_12500"))

# join data frame with the data frame extracted out of the spatial data frame
counties_swg_df <- left_join(counties_df, counties_swg) %>%
  left_join(centroids)

# plot the data

swg_yield <- c('swg_7500' = "Low yielding switchgrass", 'swg_10000' = "Medium yielding switchgrass", 'swg_12500' = "High yielding switchgrass")

ggplot(counties_swg_df) +
  geom_polygon(aes(long, lat, group = group, fill=ave_no3_leach), color = "white") +
  scale_fill_gradient(low = "white", high="#009E73", guide = "colourbar") +  
  labs(fill = expression(Average~NO[3]^{"-"}~leaching~(kg~N~ha^{-1}))) +  
  geom_point(aes(V1, V2, size = tot_no3_leach), color = "yellow") +
  labs(size = expression(Total~NO[3]^{"-"}~leaching~(Mg~N))) +
  scale_size(range = c(0, 4)) +
  coord_equal() +
  ditch_the_axes +
  facet_wrap( ~ scenario, ncol = 3, labeller = as_labeller(swg_yield))
```

## NO~3~^-^ leaching from row crops integrated with switchgrass per county
```{r plot_leach_maps,fig.width=9, fig.height=3.5, fig.cap="Figure 8: County mean annual NO~3~^-^ leaching, averaged over 2012-2015, from corn and soybean fields integrated with switchgrass."}
# set up the data frame 
counties_swg_10000 <- data.frame(id = counties$id,
                         tot_no3_leach = counties$tot_ave_no3_leach_10000,
                        ave_no3_leach = counties$ave_no3_leach_10000,
                         leach_red = counties$no3_leach_change_10000)

# join data frame with the data frame extracted out of the spatial data frame
counties_swg_10000_df <- left_join(counties_df, counties_swg) %>%
  left_join(centroids)

# plot the data



ggplot(counties_swg_10000_df) +
  geom_polygon(aes(long, lat, group = group, fill=ave_no3_leach), color = "white") +
  scale_fill_gradient(low = "white", high="#009E73", guide = "colourbar") +  
  labs(fill = expression(Average~NO[3]^{"-"}~leaching~(kg~N~ha^{-1}))) +  
  geom_point(aes(V1, V2, size = tot_no3_leach), color = "yellow") +
  labs(size = expression(Total~NO[3]^{"-"}~leaching~(Mg~N))) +
  scale_size(range = c(0, 4)) +
  coord_equal() +
  ditch_the_axes 
```

## NO~3~^-^ leaching reduction per county

```{r plot_leach_reduction_maps, fig.width=9, fig.height=3.5, fig.cap = "Figure 10: Reduction in total NO~3~^-^ leaching in the medium yield switchgrass scenario, as % of the status quo leaching in Figure 7."}


ggplot(counties_swg_10000_df) +
  geom_polygon(aes(long, lat, group = group, fill=leach_red), color = "white") +
  scale_fill_gradient(low = "white", high="#0072B2", guide = "colourbar") +  
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(NO[3]^{"-"}~leaching~reduction~("%"))) 

```



```{r shut, message=FALSE, include=FALSE}
 rm(list='isuag_db'); gc() 
```