---
title: "Modelled NO~3~ leaching in Iowa from row cropped and switchgrass fields"
author: "Elke Brandes"
date: "Tuesday, November 01, 2016"
output: html_document
---
This paper describes the data analysis of NO3 leaching results in Iowa. The initial data are the DNDC-outputs from Gabe McNunn on a clu-mapunit scale. I have performed the first analysis steps in the PostgreSQL database and access the resulting tables with the code below (using `dplyr` and `RPostgreSQL`).

##Required packages##
```{r packages, message=FALSE}
library('RPostgreSQL') # PostgreSQL database connection
library("ggplot2") # for the figures
library("maps") # to plot maps, contains country/state/county outlines
library("gridGraphics") # needed for the function unit in ggplot
library("tidyverse") # to make tidy tables, includes dplyr needed for PostgreSQL database connection
library('knitr') # to use kable() function

library(rgdal) # to read in shape files
library(ggmap)
library(scales)
library(RColorBrewer)
library("sp") # to use merge with a spatial data frame

library("tmap")


```


### Connect to the PostgreSQL datbase
With dplyr:
```{r connect_dplyr, message=FALSE}


#pw <- {
#  ""
#}
isuag_db <- src_postgres(dbname = 'isuag',
                           host = 'isu-ag-db.agsolver.com',
                           port = 5432,
                           user = 'isuag',
                           password = pw)

```
With RPostgreSQL:
```{r, connect_rpostgres, message=FALSE}
pw <- {
  "5QUXJHTbxj"
}
drv <- dbDriver("PostgreSQL") # loads the PostgreSQL driver
con <- dbConnect(drv, dbname = "isuag",  # creates a connection to the postgres database
                 host = "isu-ag-db.agsolver.com", port = 5432,
                 user = "isuag", password = pw)
# note that "con" will be used later in each connection to the database
```

### For visualization

I am using a modified theme for ggplot that has a black border and white background:
```{r theme, cache=TRUE}
theme_b_border <- function (base_size = 12, base_family = "") 
{
  theme_grey(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(size = rel(0.8), margin = margin(r=10)), #margin 
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(0.15, "cm"),
      legend.key = element_rect(colour = NA), panel.background = element_rect(fill = "white", 
      colour = NA), panel.border = element_rect(fill = NA, 
      colour = "black"), panel.grid.major = element_line(colour = NA, 
      size = 0.2), panel.grid.minor = element_line(colour = NA, 
      size = 0.5), strip.background = element_rect(fill = "grey80", 
      colour = "grey50", size = 0.2))
}
```


##1. The baseline: Profitability in corn and soybeans

### State-wide mean profitability distribution

```{r profit_distribution, fig.cap="Figure 1: Distribution of mean profitability 2012-2015. The vertical line marks the -150 US4/ha cut off value used in the profit opimization (PO) scenario for switchgrass integration below."}
profit = tbl(isuag_db, "01_profit_mean_2012_2015_aggregated") %>%
  as.data.frame()
profit$sum_ha <- profit$sum_ha * 1e-6  # convert to Mha
str(profit)

ggplot(profit, aes(profit_mean_ha_rounded, weight = sum_ha)) + 
  geom_histogram(binwidth=20, alpha = .5, position="identity") +
  theme_b_border() +
  labs(x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1})),
       y=expression(Area~(Mha))) +
  scale_x_continuous() +
  scale_y_continuous() +
  geom_vline(aes(xintercept= -150, linetype= "dashed"),  show.legend = FALSE) 
```


### Spatial distribution of highly unprofitable areas

I want to see how the areas below -150 US$/ha are distributed accross Iowa. First, I read in a county shapefile. Then, I read in the table containing the county-level unprofitable ha from the PostgreSQL database. In order to join hte data, I need to change the format of the county identifier "fips". I convert "fips" into a factor, then join the spatial and the attribute data on "fips". I then fortify (make a data frame out of) the spatial data frame and assign a column with a serial id. Finally, I join the attribute (unprofitable areas) data to this dataframe.

```{r profit_spatial, fig.cap="Figure 2: Areas below a mean 2012-2015 profitability of -150 US$/ha per county. No contraints were made with regards to subfield area size, shape or position."}
counties <- readOGR(dsn="Z:/ElkeBrandes/projects/01 subfield profit/data analysis/02 GIS/government_units/iowa_cnty/IA_cnty/county.shp",layer="county")
names(counties)[9] <- "fips" 

profit_county <- tbl(isuag_db, "07_swg_area_counties_po") %>%
  as.data.frame() 

 for (i in 1:length(profit_county$fips)) {
   profit_county$fips[i] <- sub("IA", "19", profit_county$fips[i])
    }
head(profit_county, n = 2)

# profit_county$fips <- as.factor(profit_county$fips)
counties@data <- left_join(counties@data, profit_county, by = "fips")

head(counties@data, n=2)
counties_df <- fortify(counties)
head(counties_df, n = 2)
counties$id <- row.names(counties) 
counties_profit_df <- left_join(counties_df, counties@data)

# add a spatial data frame with the county centroids, to be used to plot the total leaching N per county:
centroids <- getSpPPolygonsLabptSlots(counties) %>%
  as.data.frame() 

centroids$id <- as.character(0:98)

# join the total N leaching data to the centroids 
counties_profit_df <- left_join(counties_profit_df, centroids)

# create a theme without axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

ggplot() +
  geom_polygon(data = counties_profit_df, aes(long, lat, group = group, fill=area_in_swg_perc), color = "white") +
  geom_point(data = counties_profit_df, aes(V1, V2, size = area_in_swg), color = "orange") +
  labs(size = expression(Total~area~below~-150~"US$"~ha^{-1}~(ha))) +
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(Relative~area~below~-150~"US$"~ha^{-1}~("%"))) +
  scale_fill_gradient(low = "white", high="dark blue", guide = "colourbar") +
  ggtitle("Highly unprofitable area per county")

```



##2. Distributions of NO~3~ leaching in Iowa, from row cropped and switchgrass fields##

Connect to the isuag database and load tables that contain the rounded average NO~3~ leaching values (in kg/ha) for the **four scenarios**:

1. Status quo (row cropped areas in corn or soybean as is from 2012-2015)
2. Switchgrass low yield (all areas that were continuously in corn/soy from 2012-2015 were planted to switchgrass in 2006 for 10 years, assuming switchgrass yield is 7000 kg/ha)
3. Switchgrass medium yield (assuming a yield of 10000 kg/ha)
4. Switchgrass high yield (assuming a yield of 12500 kg/ha)

The leaching values and respective patch areas were aggregated to the same leaching values, and for each scenario, one table was created.

Use the dplyr PostgreSQL handle into the database and create a data frame:
```{r fetch_distr, cache=TRUE}

cgsb_leach <-tbl(isuag_db, "06_cgsb_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
swg_low <-tbl(isuag_db, "06_swg_7500_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
swg_med <-tbl(isuag_db, "06_swg_10000_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
swg_hi <-tbl(isuag_db, "06_swg_12500_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
```

I can run some test here:
```{r test_distr}
# send a query through dplyr
query = "SELECT SUM(area_ha) FROM \"06_cgsb_ave_no3_leach_rounded_aggr\" WHERE ave_no3_leach_cgsb_round > 60"
sum = tbl(isuag_db, sql(query))
sum
```

Create a list of data frames:
```{r list}
data_list <- list(cgsb_leach, swg_low, swg_med, swg_hi)
str(data_list[[1]])
```

Prepare a data frame to be used for the NO~3~ leaching facet plot:
```{r data, cache=TRUE}
scenario <- c(rep("cgsb", sum(sapply(data_list[1],nrow))), rep("swg_low", sum(sapply(data_list[2],nrow))), rep("swg_med", sum(sapply(data_list[3],nrow))), rep("swg_hi", sum(sapply(data_list[4],nrow)))) %>%
factor(levels= c('cgsb', 'swg_low', 'swg_med', 'swg_hi')) # make it a factor so that it is plotted in the right order in the facets
str(scenario)   

leaching <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   leaching <- append(leaching, data_list[[i]][,1])
str(leaching)  
area <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   area <- append(area, data_list[[i]][,2])
str(area)
   
leach <- data.frame(scenario, leaching, area = area*1e-6) # this factor to show Mha

head(leach)
```

Plot the distributions:
```{r histograms, cache=TRUE, fig.cap="Figure 3: Distributions of NO3 leaching in Iowa. Values were averaged over the mananagement horizon of 10 years, including two establishment years."}
scenarios <- c('cgsb'= "Status Quo (corn/soybeans)", 'swg_low' = "Low yielding switchgrass", 'swg_med' = "Medium yielding switchgrass", 'swg_hi' = "High yielding switchgrass")

ggplot(leach, aes(leaching, weight = area)) + 
  geom_histogram(binwidth=2, alpha = .5, position="identity") +
  theme_b_border() +
  labs(x=expression(Mean~NO[3]^{"-"}~leaching~2006-2015~(kg~N~ha^{-1})),
       y=expression(Area~(Mha))) +
  scale_x_continuous() +
  scale_y_continuous() +
  theme(legend.position="bottom") +
  facet_wrap( ~ scenario, ncol=2, labeller = as_labeller(scenarios)) 
```

##2. Correlation between profitability and NO3 leaching in corn and soybean fields##
I am using the table `05_dndc_clumu_cgsb_swg` in the isuag database, that contains 2012-2015 mean profit and leaching values for each of the four scenarios on a subfield (clu-map unit) scale. I want to see if there is a correlation between profitability and NO3 leaching, but since the complete data set contains more than 3 million records, I want to randomly select a sub-sample for the correlation.

Using a SQL chunk, create a table from a sub-sample (n=10000) *within* the PostgreSQL database. Setseed is used so that every time I run the code I will get the same random sample.
```{sql, connection=con, message=FALSE, cache=TRUE}
SELECT setseed(0.5)
```
```{sql, connection=con, message=FALSE, cache=TRUE}
DROP TABLE IF EXISTS "r_dndc_clumu_cgsb_swg_sample";
```
```{sql, connection=con, message=FALSE, cache=TRUE}
CREATE TABLE "r_dndc_clumu_cgsb_swg_sample" AS SELECT mean_profit_ha, ave_no3_leach_ha_cgsb FROM "05_dndc_clumu_cgsb_swg" ORDER BY random() LIMIT 10000
```


Pull the new table into R:
```{r data_sample}
sample = tbl(isuag_db, "r_dndc_clumu_cgsb_swg_sample") %>%
  as.data.frame(sample)

head(sample)
str(sample)
```

Plot the sample:
```{r scatterplot, cache=TRUE, fig.cap="Figure 4: Scatterplot of a subsample (n=10,000) of profitability and corn/soy leaching data on subfield areas in Iowa. The red lines indicate possible cut off values, such as -150 US$ ha^-1^ for profitability (solid) and 60 kg N ha^-1^ for NO3 leaching (dashed)."}
ggplot() +
  geom_point(data = sample, aes(x = mean_profit_ha, y = ave_no3_leach_ha_cgsb)) +
  theme_b_border() +
  labs(y=expression(Mean~NO[3]^{"-"}~leaching~2006-2015~(kg~N~ha^{-1})),
       x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1}))) +
  geom_vline(aes(xintercept= -150, linetype= "dashed", color = "red"),  show.legend = FALSE) +
  geom_hline(aes(yintercept= 60, linetype= "dotted", color = "red"),  show.legend = FALSE)
  
```

There are many records with leaching values close to 0, and then only a few that are higher but below about 10. The next plot zooms in on the y axis to see this better:
```{r scatterplot_zoom, cache=TRUE, fig.cap="Figure 5: Scatterplot of a subsample (n=10000) of profitability and corn/soy leaching data on subfield areas in Iowa. The y axis is zoomed in on 0-20 kg N ha^-1^ to better see the low leaching values. The red line indicates a possible cut off value of -150 US$ ha^-1^ for profitability."}
ggplot() +
  geom_point(data = sample, aes(x = mean_profit_ha, y = ave_no3_leach_ha_cgsb)) +
  theme_b_border() +
  labs(y=expression(Mean~NO[3]^{"-"}~leaching~2006-2015~(kg~N~ha^{-1})),
       x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1}))) +
  scale_x_continuous() +
  scale_y_continuous(limits=c(0,20)) +
  geom_vline(aes(xintercept= -150, linetype= "dashed", color = "red"),  show.legend = FALSE)
  
```
In order to know how much of the area is included in those very low leaching patches, I run queries in the SQL database:
```{sql, connection=con, message=FALSE, cache=TRUE}
SELECT sum(clumuha) FROM "05_dndc_clumu_cgsb_swg" WHERE ave_no3_leach_ha_cgsb < 0.1
```
```{sql, connection=con, message=FALSE, cache=TRUE}
SELECT sum(clumuha) FROM "05_dndc_clumu_cgsb_swg" WHERE ave_no3_leach_ha_cgsb < 0.5
```
```{sql, connection=con, message=FALSE, cache=TRUE}
SELECT sum(clumuha) FROM "05_dndc_clumu_cgsb_swg" WHERE ave_no3_leach_ha_cgsb < 1
```
```{sql, connection=con, message=FALSE, cache=TRUE}
SELECT sum(clumuha) FROM "05_dndc_clumu_cgsb_swg" WHERE ave_no3_leach_ha_cgsb > 1 AND ave_no3_leach_ha_cgsb < 10
```

##3. Switchgrass integration scenarios

From figure 2 we can see that the relation between NO3 leaching and profitability is not completely random. At higher profitability, the number of sub-field patches with high NO3 leaching values decreases. The lower the profitability, the higher the heterogeneity in leaching values. In other words, by managing the areas of lowest profitability in switchgrass would target those areas of highest leaching values (but also many of lower leaching values). Managing the areas of highest leaching (e.g. above a cut off of 100 kg N/ha) would mostly affect the areas that were on average unprofitable between 2012 and 2015.

I am choosing -250 US$/ha as a profitability cut off and 100 kg N/ha as an NO3 leaching cut off. That means, I produce two switchgrass integration scenarios:

1. Profit-optimized (PO): Switchgrass is grown on the sub-field areas where mean profitability is below -150 US$/ha
2. Water quality-optimized (WO): Switchgrass is grown on the sub-field areas where NO3 leaching is above 60 kg N/ha

Each of those scenario has three "sub-scenarios" assuming the three different switchgrass yields, so I end up having 6 scenarios that I want to compare to the status quo.

Comparison criteria:

Improvement of total N leaching in form of NO3 state wide and on the county level (in % of status quo values).

Leaching sums are calculated in PostgreSQL. Access the results:
```{r total_leaching}
leach_tot <-tbl(isuag_db, "05_dndc_no3_leach_sums_iowa_scenarios")  %>%
  as.data.frame()
names(leach_tot) <- c("Status Quo", "PO 7500", "PO 10000", "PO 12500","WO 7500", "WO 10000", "WO 12500")

leach_red <- data.frame(status_quo = "NA",
                        po_7500 = (leach_tot[1,1]-leach_tot[1,2])*100/leach_tot[1,1],
                        po_10000 = (leach_tot[1,1]-leach_tot[1,3])*100/leach_tot[1,1],
                        po_12500 = (leach_tot[1,1]-leach_tot[1,4])*100/leach_tot[1,1],
                        wo_7500 = (leach_tot[1,1]-leach_tot[1,5])*100/leach_tot[1,1],
                        wo_10000 = (leach_tot[1,1]-leach_tot[1,6])*100/leach_tot[1,1],
                        wo_12500 = (leach_tot[1,1]-leach_tot[1,7])*100/leach_tot[1,1])
names(leach_red) <- c("Status Quo", "PO 7500", "PO 10000", "PO 12500","WO 7500", "WO 10000", "WO 12500")

leach_table <- rbind(leach_tot[1,],leach_red[1,])

kable(leach_table, col.names=c("Status Quo","PO 7500", "PO 10000", "PO 12500","WO 7500", "WO 10000", "WO 12500"), caption="Table 1: Annual NO3 leaching (Mg N) in Iowa modelled with DNDC under different switchgrass integration scenarios. PO: Profitability optimized, WO: Water quality optimized." )

```


##4. Disproportionate benefits?

I want to see how NO3 leaching reduction increases with increasing area in switchgrass, to test if the disproportionate benefits hypothesis holds true in our model.

I want to select the total area of polygons that fall under specific profitability thresholds (PO scenario).
```{r swg_areas_po, cache=TRUE}
swg_areas_po <- vector(mode="numeric", length=0)
# profit <- tbl(isuag_db, "05_dndc_clumu_cgsb_swg")
profit_cutoffs <- c(-500, -400, -300, -200, -150, -100, -50, 0, 50, 100, 150)
 for (i in 1:length(profit_cutoffs)) {
    query <- paste("SELECT SUM(clumuha) FROM \"05_dndc_clumu_cgsb_swg\" WHERE mean_profit_ha < ", as.character(profit_cutoffs[i]))
    sum_area <- tbl(isuag_db, sql(query)) %>%
      as.data.frame()
    sum_area <- sum_area[1,1]
    swg_areas_po <- append(swg_areas_po, sum_area) 
  }
```

```{r no3_leaching_sums, cache=TRUE}
sum_leaching <- tbl(isuag_db, "05_dndc_no3_leach_sums_iowa_dispr_benefits") %>%
  as.data.frame() %>%
  gather(corn_soybeans, swg_7500_1, swg_7500_2, swg_7500_3, swg_7500_4, swg_7500_5, swg_7500_6, swg_7500_7, swg_7500_8, swg_7500_9, swg_7500_10, swg_7500_11, swg_10000_1, swg_10000_2, swg_10000_3, swg_10000_4, swg_10000_5, swg_10000_6, swg_10000_7, swg_10000_8, swg_10000_9, swg_10000_10, swg_10000_11, swg_12500_1, swg_12500_2, swg_12500_3, swg_12500_4, swg_12500_5, swg_12500_6, swg_12500_7, swg_12500_8, swg_12500_9, swg_12500_10, swg_12500_11, key = swg_scenarios, value=ave_no2_leach)
```

```{sql connection=con, message=FALSE, cache=TRUE}
select sum(clumuha) from "05_dndc_clumu_cgsb_swg";
```

I want to select the total area of polygons that fall above specific N leaching thresholds (WO scenario).
```{r swg_areas_wo, cache=TRUE}
swg_areas_wo <- vector(mode="numeric", length=0)
leach_cutoff <- c(100, 90, 80, 70, 60, 50)
 for (i in 1:length(leach_cutoff)) {
    query <- paste("SELECT SUM(clumuha) FROM \"05_dndc_clumu_cgsb_swg\" WHERE ave_no3_leach_ha_cgsb > ", as.character(leach_cutoff[i]))
    sum_area <- tbl(isuag_db, sql(query)) %>%
      as.data.frame()
    sum_area <- sum_area[1,1]
    swg_areas_wo <- append(swg_areas_wo, sum_area) 
  }
```

Create data frames for figures:
```{r data_dispr_benefits}
swg_yield <- as.factor(c("NA",rep("7500",11), rep("10000",11),rep("12500",11)))
profit_cutoff <- c(NA, rep(profit_cutoffs,3))
swg_area_po <- c(0,rep(swg_areas_po,3))
swg_area_po_percent <- (swg_area_po/9374362)*100
N_leaching <- sum_leaching[,2]
N_reduction <- abs((N_leaching - N_leaching[1])/N_leaching[1])*100

relation_po <- data.frame(swg_yield, profit_cutoff, swg_area_po_percent, N_reduction)

swg_area_wo_percent <- (swg_areas_wo/9374362)*100

relation_wo <- data.frame(leach_cutoff, swg_area_wo_percent)
```

Plot relationsships between thresholds and % area in switchgrass:
```{r thresholds, fig.show="hold", fig.height= 3, fig.width=4.5, fig.cap="Figure 6: Relation between the set thresholds in the profitability (left panel) and water quality optimized (right panel) scenario and the resulting area in switchgrass."}
ggplot() +
  geom_point(data = relation_po, aes(x = profit_cutoff, y = swg_area_po_percent)) +
  theme_b_border() +
  labs(x=expression(Profitability~threshold~("US$"~ha^{-1})),
       y=expression(Area~"in"~switchgrass~("%"))) +
  scale_x_continuous(breaks=seq(from=-500, to=200, by= 100)) +
  scale_y_continuous(limits= c(0,65)) +
    geom_vline(aes(xintercept= -150, linetype= "dotted", color = "red"),  show.legend = FALSE)

ggplot() +
  geom_point(data = relation_wo, aes(x = leach_cutoff, y = swg_area_wo_percent)) +
  theme_b_border() +
    labs(x=expression(NO[3]^{"-"}~leaching~threshold~(kg~ha^{-1})),
       y=expression(Area~"in"~switchgrass~("%"))) +
  scale_x_reverse(breaks=seq(from=50, to=100, by= 10)) +
  scale_y_continuous(limits= c(0,65)) +
    geom_vline(aes(xintercept= 60, linetype= "dotted", color = "red"),  show.legend = FALSE)


```



Plot relationships:
```{r relations, cache=TRUE, fig.show="hold", fig.height= 3, fig.width=4.5, fig.cap="Figure 7: Increase in NO~3~ leaching reduction with increasing area in switchgrass (left panel), and with increasing profitability threshold (right panel)."}

ggplot() +
  geom_point(data = relation_po, aes(x = swg_area_po_percent, y = N_reduction, color=swg_yield)) +
  theme_b_border() +
  labs(x=expression(Area~"in"~switchgrass~("%")),
       y=expression(NO[3]^{"-"}~leaching~reduction~("%"~N))) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete(name="Switchgrass\nYield (kg/ha)",  guide = FALSE)

ggplot() +
  geom_point(data = relation_po, aes(x = profit_cutoff, y = N_reduction, color=swg_yield)) +
  theme_b_border() +
  labs(x=expression(Profitability~threshold~("US$"~ha^{-1})),
       y=expression(NO[3]^{"-"}~leaching~reduction~("%"~N))) +
  scale_x_continuous(breaks=seq(from=-500, to=200, by= 100)) +
  scale_y_continuous() +
  scale_color_discrete(name="Switchgrass\nYield") +
  theme(legend.position=c(0.2,0.7))
```

##5. Spatial distribution of NO~3~ leaching changes##

Read in the table from the isuag database that contains county identifiers ("fips") and leaching results. 
```{r counties_attributes}
 counties_leach = tbl(isuag_db, "07_dndc_counties") %>%
   as.data.frame()
 # change the data in the column named "FIPS", so that is contains the county identifiers in the same format as the spatial file:
 for (i in 1:length(counties_leach$fips)) {
   counties_leach$fips[i] <- sub("IA", "19", counties_leach$fips[i])
    }
 # make it a factor
  counties_leach$fips <- as.factor(counties_leach$fips)

head(counties_leach, n=2)
```

Read in a shape file with county delineations. I could use a map provided in the maps package or read in a file from my directory, which I do in this case:
```{r counties_spatial}

counties <- readOGR(dsn="Z:/ElkeBrandes/projects/01 subfield profit/data analysis/02 GIS/government_units/iowa_cnty/IA_cnty/county.shp",layer="county")

names(counties)[9] <- "fips" # change to lower case so that we can join on this field later on
summary(counties)

head(counties@data, n=2) # shows only the slot "data"
head(counties@polygons[[1]]@Polygons[[1]]@coords, 3) # shows the coordinates

# join the attribute data with the spatial data, using a left join and the identifier "fips":
counties@data <- left_join(counties@data, counties_leach, by = "fips")
```
I can map the attributes using tmap, but I want to use ggplot. Therefore I am not using the following code:
```qtm(counties, "ave_no3_leach_cgsb")```

To plot spatial data in ggplot, however, I need to make a data frame out of the SpatialPolygonDataFrame to get access to the longitude and latitude. This data frame contains a row for each vertex that the polygons are made of, and an ID starting with 0 up to 98. The attribute data will be lost in this data frame. Therefore, it needs to be joined back afterwards.
```{r join} 
counties_df <- fortify(counties)
head(counties_df, n = 2)
counties$id <- row.names(counties) # assign a column with the variable ID that is equal to the row names (=numbers), starting with 0.
head(counties@data, n = 2)
# join the attributes back to the data frame, to create a dataframe used to map the baseline (corn/soybean):
counties_baseline_df <- left_join(counties_df, counties@data)
```

```{r plot_baseline_map, fig.cap="Figure 8: County mean annual NO~3~ leaching, averaged over 2012-2015, from corn and soybean fields in Iowa, are shown as a color scale. Bubbles indicate the total N that has been leaching on average per year from each county."} 

head(counties_baseline_df, n= 2)

# create a theme without axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )


# add a spatial data frame with the county centroids, to be used to plot the total leaching N per county:
centroids <- getSpPPolygonsLabptSlots(counties) %>%
  as.data.frame() 
centroids$id <- as.character(0:98)

# join the total N leaching data to the centroids 
counties_baseline_df <- left_join(counties_baseline_df, centroids)


ggplot() +
  geom_polygon(data = counties_baseline_df, aes(long, lat, group = group, fill=ave_no3_leach_cgsb), color = "white") +
  geom_point(data = counties_baseline_df, aes(V1, V2, size = tot_ave_no3_leach_cgsb), color = "yellow") +
  labs(size = expression(Total~NO[3]^{"-"}~leaching~(Mg~N~ha^{-1}))) +
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(Average~NO[3]^{"-"}~leaching~(kg~N~ha^{-1}))) +
  scale_fill_gradient(low = "white", high="#009E73", guide = "colourbar") +
  ggtitle("Average and total annual NO3 leaching per county")

head(counties_baseline_df, n=2)


```

Now, I want to plot two times three maps in a facet plot that show the NO~3~ leaching results and the leaching reduction, when switchgrass is grown on areas losing -150 US$/ha or more, for each of the three yield scenarios. For the facet plot in ggplot, I need to re-arrange the data, creating the variable "scenario" and adding two columns with values, one for the average leaching and one for the percent change from the baseline to the switchgrass integration scenario.
```{r plot_leach_maps,fig.width=10, fig.height=3, fig.cap="Figure 9: County mean annual NO~3~ leaching, averaged over 2012-2015, from three switchgrass integration scenarios in Iowa, are shown as a color scale. Bubbles indicate the total N that has been leaching on average per year from each county. Switchgrass was planted on areas that lost -150 US$/ha or more on average between 2012 and 2015. The three maps refer to different yield assumptions for switchgrass (low yielding: 7,500 kg/ha, medium yielding: 10,000 kg/ha, and high yielding: 12,500 kg/ha)"}
# set up the data frame 
counties_swg <- data.frame(id = rep(counties$id,3),
                         scenario = c(rep("swg_7500", length(counties$id)), rep("swg_10000", length(counties$id)), 
                                                                               rep("swg_12500", length(counties))),
                         tot_no3_leach = c(counties$tot_ave_no3_leach_7500, counties$tot_ave_no3_leach_10000,
                                           counties$tot_ave_no3_leach_12500),
                         ave_no3_leach = c(counties$ave_no3_leach_7500, counties$ave_no3_leach_10000, counties$ave_no3_leach_12500),
                         leach_red = c(counties$no3_leach_change_7500, counties$no3_leach_change_10000,
                                       counties$no3_leach_change_12500))

counties_swg$scenario <- factor(counties_swg$scenario, levels = c("swg_7500", "swg_10000", "swg_12500"))
head(counties_swg, n = 2)
# join data frame with the data frame extracted out of the spatial data frame
counties_swg_df <- left_join(counties_df, counties_swg) %>%
  left_join(centroids)
head(counties_swg_df, n = 4)

# plot the data

swg_yield <- c('swg_7500' = "Low yielding switchgrass", 'swg_10000' = "Medium yielding switchgrass", 'swg_12500' = "High yielding switchgrass")

ggplot(counties_swg_df) +
  geom_polygon(aes(long, lat, group = group, fill=ave_no3_leach), color = "white") +
  scale_fill_gradient(low = "white", high="#009E73", guide = "colourbar") +  
  labs(fill = expression(Average~NO[3]^{"-"}~leaching~(kg~N~ha^{-1}))) +  
  geom_point(aes(V1, V2, size = tot_no3_leach), color = "yellow") +
  labs(size = expression(Total~NO[3]^{"-"}~leaching~(Mg~N))) +
  scale_size(range = c(0, 4)) +
  coord_equal() +
  ditch_the_axes +
  facet_wrap( ~ scenario, ncol = 3, labeller = as_labeller(swg_yield)) +
  ggtitle("Average and total annual NO3 leaching per county")
```

```{r plot_leach_reduction_maps, fig.width=10, fig.height=3, fig.cap = "Figure 10: Reduction in total NO~3~ leaching in the switchgrass integration scenarios, as % of the status quo leaching in Figure 7."}
ggplot(counties_swg_df) +
  geom_polygon(aes(long, lat, group = group, fill=leach_red), color = "white") +
  scale_fill_gradient(low = "white", high="#0072B2", guide = "colourbar") +  
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(NO[3]^{"-"}~leaching~reduction~("%"))) +
  facet_wrap( ~ scenario, ncol = 3, labeller = as_labeller(swg_yield))

```


Shut the database down:
```{r shut, message=FALSE}
 rm(list='isuag_db'); gc() 
```
