---
title: Addressing farm economy, water quality, and bioenergy feedstock production by sub-field switchgrass integration on unprofitable cropland
author: "Elke Brandes"
date: "Friday, January 20, 2017"
output:
  word_document: default

bibliography: nutrient_bibliography.bib
csl: environmental-research-letters.csl

---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning = FALSE)

```
## Authors

Elke Brandes

Gabe McNunn

David Muth

Lisa Schulte

Andy VanLoocke

Emily Heaton


## Abstract


Progress on reducing nutrient leaching from Iowa cropland has been slow due to perceived conflicts between short-term profitability and long-term land stewardship. Besides mitigating nitrate leaching from cropland, perennial biomass crops such as switchgrass can improve short- and long-term farm level profitability and help meeting national bioenergy targets.

To explore the potential of socially compatible switchgrass integration into Iowa's agricultural landscape, we combined two subfield-scale data sets: (i) a profitability analysis and (ii) nitrate leaching rates estimated with the DeNitrification-DeComposition (DNDC) model. We based the management change decision from row crops to switchgrass on status-quo (2012-2015) profitability and nitrate leaching rates, creating a matrix of possible threshold combinations. For each scenario resulting from these combinations we calculated: 1) the percent of Iowa cropland in switchgrass, 2) resulting biomass production, and 3) the relative reduction in nitrate leaching compared to row crops. 

Converting all subfield areas losing > US\$ 100 ha-1 yr1 and leaching > 50 kg N ha-1 yr-1 to switchgrass would result in 18% nitrate leaching reduction for Iowa. We found that the counties of highest status-quo nitrate leaching did not line up with those of highest reduction potential. *Within the impaired Upper Raccoon River Watershed, converting all subfield areas losing > US\$ 100 ha-1 yr1 and leaching > 50 kg N ha-1 yr-1 to switchgrass resulted in 11.2% of row cropland converted to switchgrass and 8% nitrate leaching reduction. (needs to be updated)* Most subfield areas converted to switchgrass displayed high nitrate leaching reduction (> 50%), but the high productivity of most farmland in this watershed resulted in a relatively small conversion area and therefore a low mean nitrate leaching reduction. Our analysis disaggregates previous approaches of county- or watershed-scale nitrogen balance down to a finer resolution and takes into account the biogeochemical processes driving nitrogen loss in agroecosystems. Even though approximated, our analysis can be turned into a farm-level decision making tool to identify non-linear responses of nitrate leaching with the strategic integration of perennial vegetation.

## Highlights 
(not needed for ERL, but can be useful anyway)

- In a three way comparison we analyse the impact of cropland conversion to a dedicated biomass crop on nutrient loss improvement
- We modelled the leachable N fraction for agricultural lands in the state of Iowa
- using high-resolution spatial data to test ecosystem service outcomes of a socio-economic approach of perennial biomass crop integration

## Introduction

### Harnessing economic interests for switchgrass integration

Pooled under the term precision conservation, efforts have been made recently to develop landscape designs that meet agricultural needs while optimizing environmental risk mitigation by targeted placement of conservation management practices [@mcconnell2011precision]. This approach can potentially be adapted to target perennial crops for sustainable cellulosic bioenergy production [@chaubey2016precision]. Due to their multiple benefits on water and soil quality, perennial biomass crops have been considered suitable for conservation purposes while providing feedstock for a growing bioeconomy [@tilman2009beneficial; @langholtz2016billion]. However, in the past decade, large extents of CRP and grassland have been converted back to cropland [@morefield2016grasslands; @lark2015cropland]. In practice, perceived socio-economic barriers prevent the implementation of urgently needed conservation management [@james2010profitability]. However, despite this seemingly universal trade-off between economic and environmental desires, there are indications of synergies between economic and environmental risk mitigation: within a single corn field, @ssegane2015multifunctional showed that areas of low crop yield coincided with high NO~3~^-^ leaching due to low soil fertility related to high water infiltration.

*add a sentence proving evidence to support we are working in a key niche with the sub-field scale: Are there management recommendations?* 

### Water quality as the ecosystem service of interest

Nitrate, the nitrogen form that is readily available to crops, is also subject to leaching through the root zone and drainage tiles into surface waters. Non-point nutrient pollution from industrialized agriculture in the Mississippi River Basin has been exacerbating the hypoxic zone in the Gulf of Mexico [@turner2003linking]. Over half of the  nitrogen delivered through this route originate from cropland cultivated with maize and soybeans [@alexander2008differences]. @schilling2000relationship showed a strong relationship between percent land area in row crop and nitrate-N concentration in watersheds in Iowa. Their study suggests a 0.1 mg l^-1^ reduction of nitrate-N in the surface water with every percent reduction of row crop area in the above watershed.

The Mississippi River Gulf of Mexico Watershed Nutrient Task Force [-@river2008action] has set the goal to decrease the hypoxic zone's 5 year average to under 5000 km^2^. The @inrs2013 translates this goal to a 45% nitrogen reduction requirement for Iowa and recommends a wide range of nitrogen conservation practices including perennial crops. However, the spatial heterogeneity on a level relevant for decisionmaking is not accounted for in this strategy.

Perennial crops can drastically improve nitrate losses compared to row crops. Several experiments growing maize-soybean rotations side by side with perennial grasses such as Miscanthus x giganteus, switchgrass, or reconstructed prairie, resulted in leaching reductions of more than 90% in the perennial plots [@mcisaac2010miscanthus; @smith2013reduced; @daigh2015subsurface]. These experiments were conducted on fertile soils, but large uncertainties exist on how perennials compare to row crops on less productive cropland, where they would be most likely grown to provide feedstock to an emerging bioeconomy. Moreover, the extent and strategic location of perennials in the agricultural landscape effective to reach the nutrient reduction goal remains unknown.

@vanloocke2016din found that replacing maize currently grown for bioethanol with cellulosic biomass crops in the Atchafalaya and Mississippi River Basin could reduce dissolved inorganic nitrogen (DIN) export to the Gulf of Mexico by 15 to 20%. (add sentence citing Davis et al 2011 or 2012).They suggest that integration of perennial grasses targeted to areas of low row crop productivity and high leaching rates could further increase surface water quality. Many modelling approaches have been conducted on a comparably coarse scale. These water quality studies of larger regions fail to reflect the high-resolution farm units relevant for decision making.
As part of a nutrient budget for Iowa, @christianson2012nitrogen recently estimated nitrate leaching from fertilizer inputs. However, their approach did not take into account weather and soil related variability.

Not only environmental considerations play a role in strategic placement of perennial crops, but also economic constraints. Areas of high environmental risk are assumed to overlap with areas of high economic risk - targeting those areas may provide opportunities to combine financial risk mitigation with conservation management [@muth2014profitability]. We argue that targeted replacement of row crops with the highly productive and locally adapted switchgrass variety "Liberty" on low-performing cropland areas could contribute to a reduction of N loss in Iowa while enhancing biomass production and further regulating ecosystem services.

To estimate the potential of nitrate leaching reduction by switchgrass integration in a spatial resolution relevant for decision makers, we apply a mechanistic model that accounts for weather and soil dependent processes of the nitrogen cycle to a subfield scale across Iowa. As an example for an impaired watershed, the North Racoon River Watershed is in focus in this paper to shed light on potential mitigation efforts and their effectiveness.

We are presenting a novel approach to precision conservation and perennial integration: placing switchgrass on crop land that is least socially desirable to grow row crops, using a three-way comparison:
   
- baseline scenario: cropland in maize and /or soybean between 2012 and 2015
- minimum (tweak) scenario: mitigating profit loss and nitrate leaching while keeping the area in switchgrass at around 10%
- nutrient reduction scenario: achieving a 40% reduction in nitrate leaching compared to the status quo, thereby achieving the Iowa nutrient reduction goal mainly with dedicated biomass production

With these scenarios located along a wide range of landscape integration schemes, we aim to identify areas in Iowa with highest potential for switchgrass integration that is compatible with socio-economic constraints, and to quantify the extent of nitrate leaching reduction by converting cropland areas to switchgrass that are at highest economic and environmental risk.

## Methods

Using publicly available data, we identified subfield cropland areas of low profitability. We estimated nitrate leaching with the processed-based DeNitrifiction-DeComposition model (DNDC) for a baseline, including all cropland that was continuously in maize and/or soybeans 2012-2015, and three switchgrass scenarios of varying biomass yields. We tested the sensitivity of three model outcomes for Iowa (% area in switchgrass, total biomass production, and relative nitrate leaching reduction) for the management change decision points for nitrate leaching and profitability.



### Spatial data

We included all Iowa cropland that was managed in either maize or soybeand continuously between 2012 and 2015. Subfield units were created by intersecting the latest available common land use (CLU) boundaries [@usda2008clu] with the Soil Survey map unit delineations [@nrcs2016soil]. Crop cover for each year was identified for each CLU polygon from the respective cropland data layer (CDL) raster files [@nass2016cdl] as described in @gelder2007land.

### Economic analysis

We performed a profitability analysis on a sub-field resolution following the methods in @brandes2016subfield, but updated the considered time frame to 2012-2015. For each sub-field polygon, defined by field (CLU) and soil characteristics (Soil Survey), we estimated the profitability, taking into account the management type (maize following maize, maize following soybeans, and soybeans), the respective crop production costs, cash rents, yields, and grain prices for each year.
We derived potential crop yields from the Corn Suitability Rating (CSR2), an index specific to Iowa that integrates a number of soil characteristic and an expert judgement correction factor  [@burras2015csr2], using a linear function [@sassman2015ispaid]. To include spatial and temporal yield variability, we scaled the potential yieds to county averages reported by the National Agricultural Statistics Service [@nass2016stats]. To estimate field-scale cash rents, we scaled county averages reported in annual surveys performed by Iowa State University Extension and Outreach [@plastina2016rents] to CSR2 (sub-field scale) and then calculated field averages [see details in @brandes2016subfield].

### DNDC model description

DNDC is a process-based, field-scale biogeochemistry model for agricultural systems that estimates the mass balances of nitrogen and carbon on a daily resolution. Processes are simulated in the six sub-models soil climate, plant growth, decomposition, denitrfication, nitrification, and fermentation. The main driving inputs into the model are climate, soil, vegetation, and management practices. An overview of the model is described in [@tonitto2010application]. Since the study aim is focussing on water quality, we extracted the model results of nitrate leaching as well as ammonium volatilization, an output that is closely connected to leaching through the microbial process of nitrification.

- we performed four model runs for Iowa: 
    + baseline (continuous maize or maize-soybean rotation) 
    + three runs for switchgrass, assuming high (12,500 kg ha^-1^), medium (10,000 kg ha^-1^), and low (7,500 kg ha^-1^) yields
- in the switchgrass runs, all fields considered in the baseline run were assumend to be managed in switchgrass.
- data used: *to be completed with information from Gabe*
    + climate data: county-level daily min/maz temperatures and precipitation from Daymet [@thornton2016daymet]
    + soil data: National Soil Survey Geographic Database [SSURGO, @nrcs2016soil] 
    + land use and management: adapted to common practices in Iowa (for row crops) and recommendations for switchgrass *(add table with numbers)*
    + crop parameters *(add table with numbers)*

### State-wide switchgrass integration scenarios
In the management decision framework presented in this study, sub-field areas that are most economically and environmentally unfavorable in the baseline scenario are targeted for switchgrass integration. In a first step, we examine the state-wide model outcomes for i) relative cropland area converted to switchgrass, ii) biomass production from switchgrass, and iii) relative nitrate leaching reduction, as a function of both profitability and nitrate leaching cut-off decision points. Since the state-wide model outcomes showed a low sensitity to switchgrass yield (Figure S4), we performed further analysis with the medium switchgrass yield of 10,000 kg ha^-1^.

We applied every possible combination of profitability and nitrate leaching cut-off, covering the range of 2012-2015 mean profitability (US\$-800 ha^-1^ to US\$ 700 ha^-1^, by steps of US\$ 100 ha^-1^) as well as the range of 2012-2015 mean nitrate leaching (0 - 150 kg ha^-1^, by steps of 10 kg ha^-1^)
and analyze their outcomes on the state level, resulting in 256 (16 x 16) distict scenarios.

### Closer look at select scenarios
From these scenarions, we selected two cases based on their state-wide outcomes to analyze in further detail:
1. Tweak scenario: Convert all cropland with profitability of < US\$ -100 ha^-1^ and nitrate leaching of > 50 kg N ha^-1^. This scenario was driven by the goal to convert approx. 10% of cropland area to switchgrass.
2. Nutrient reduction scenario: Convert all unprofitable cropland (with profitability of < US\$ 0 ha^-1^) that loses > 20 kg N ha^-1^ through nitrate leaching. This scenario was driven by the goal to reach a substantial portion of the nitrate leaching reduction goal (45\%) set in the Iowa Nutrient Reduction Strategy.

For these two scenarios, we compared leaching results to the baseline on a county resolution to identify counties with high potential for water quality improvement. We also studied the impared upper part of the North Raccoon River Watershed to relate opportunities for water quality improvement with perennial integration as to the rest of Iowa.

## Results

- *create a scatter plot showing county mean annual N leaching as a function of county precip date used in the model. That will show the annual variation in leaching that is hidden in the four year averages.*

- Mean profitability shows spatial heterogeneity accross Iowa, with highest total area losing > US$ 100 ha^-1^ in west central Iowa and highest relative area in the west central and south part of the state (Figure 1)

- Surface plots and contour plots show the resulting area in switchgrass for Iowa, with varying combined thresholds of profitability and nitrate leaching (including cropland below a cut-off profitability and above a cut-off nitrage leaching) (Figure 2, 3)

- selected criteria for conversion to switchgrass: 

  + tweak scenario: mean profitability < US$ -100 ha^-1^, N loss > 50 kg ha^-1^ -- resulting in 11.4% in switchgrass, 10.7 million Tons biomass, and 18% nitrate leaching reduction (Figure 3)
  + nutrient reduction scenarion: mean profitability < US$ 0 ha^-1^, N loss > 20 kg ha^-1^ -- resulting in 35% in switchgrass, 32.8 million Tons biomass, and 37.9% nitrate leaching reduction (Figure 3)

- State-wide mean annual N loss (status quo, 2012-2015) amounted to 0.44 million metric tonnes. On the state-wide average, converting all areas with profitability below US$ -100 ha^-1^ and with N losses of > 50 kg ha^-1^ to switchgrass would result in a 17.3, 18.1, and 18.6 % reduction of N loss for the low, medium, and high yielding scenario (Table 1) 

- the spatial pattern of nitrate leaching in Iowa shows highest leaching in the counties on the Des Moines Lobe and in the north east of the state (Figure 4)

- according to the model presented here, highest nitrate leaching reduction with the switchgrass integration scenario can be achieved in some of the central and eastern counties (Figure 5). The counties of highest nitrate leaching in the status quo scenario do not line up with those of highest reduction potential.

- the subfield resolution map of nitrate leach reduction (Figure 6) illustrates the spatial heterogeneity of the environmental outcomes of the biomass crop integration. 

- Strong "county effects" seem to drive the highest reduction values (Figure 6(c)). Some counties with very high conversion areas (Figure 1) and therefore high potential nitrate leaching reduction (Figure 4, 5), e.g., Monona, Carroll, and Hamilton County, both show a low yield/cash rent ratio (data not shown).

- zooming in on a select impaired watershed in Iowa: Pattern of switchgrass fields vary, some regions, whole fields are converted, in other, subfield areas are patchy and more fragmented (Figure 6(c)). In the North Racoon River Watershed, an area of high (80%) row crop percentage, the tweak scenario (converting cropland that loses > US$ 100 and > 50 kg N ha^-1^ to switchgrass) would result in 14.3 % of row crops converted to switchgrass and a nitrate leaching reduction of 11.3 %. Most subfield areas converted to switchgrass display nitrate leaching reduction of over 50%, but the relative small conversion area (due to relative high profitability in this watershed) produces a low mean nitrate leaching reduction. The nutrient reduction scenario (converting all unprofitable land that is leaching > 20 kg N ha^-1^ to switchgrass) would result in 39.7% of row crops converted to switchgrass and a nitrate leaching reduction of 31.4%.


```{r packages}
library('RPostgreSQL') # PostgreSQL database connection
library("ggplot2") # for the figures
library("maps") # to plot maps, contains country/state/county outlines
library("gridGraphics") # needed for the function unit in ggplot
library("tidyverse") # to make tidy tables, includes dplyr needed for PostgreSQL database connection
library('knitr') # to use kable() function

library(rgdal) # to read in shape files
library(ggmap)
library(scales)
library(RColorBrewer)
library("sp") # to use merge with a spatial data frame

library("tmap")

library(png) # to read png files
library(grid) # to display png files
library("plot3D") # for surface plot

library('matrixStats')
library(gridExtra) # need to plot ggplots in a grid

library('foreign')


```

```{r connect_dplyr}

# with dplyr:
pw <- "****" # enter password here
HeatonAgsolver_db <- src_postgres(dbname = 'AgSolver',
                           host = 'heatonagsolver.cv5vlzyvdngx.us-east-1.rds.amazonaws.com',
                           port = 5432,
                           user = 'heatonlab',
                           password = pw)
```

```{r, connect_rpostgres}

# With RPostgreSQL:
drv <- dbDriver("PostgreSQL") # loads the PostgreSQL driver
con <- dbConnect(drv, dbname = "AgSolver",  # creates a connection to the postgres database
                 host = "heatonagsolver.cv5vlzyvdngx.us-east-1.rds.amazonaws.com", port = 5432,
                 user = "heatonlab", password = pw)
# note that "con" will be used later in each connection to the database
```

```{r theme, cache=TRUE}
# I am using a modified theme for ggplot that has a black border and white background:
theme_b_border <- function (base_size = 12, base_family = "") 
{
  theme_grey(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(size = rel(0.8), margin = margin(r=10)), #margin 
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(0.15, "cm"),
      legend.key = element_rect(colour = NA), panel.background = element_rect(fill = "white", 
      colour = NA), panel.border = element_rect(fill = NA, 
      colour = "black"), panel.grid.major = element_line(colour = NA, 
      size = 0.2), panel.grid.minor = element_line(colour = NA, 
      size = 0.5), strip.background = element_rect(fill = "grey80", 
      colour = "grey50", size = 0.2))
}

# I am using a modified filled contour plot function for the figure that shows switchgrass area as a function of cut off values:

filled.contour3 <-
  function (x = seq(0, 1, length.out = nrow(z)),
            y = seq(0, 1, length.out = ncol(z)), z, xlim = range(x, finite = TRUE), 
            ylim = range(y, finite = TRUE), zlim = range(z, finite = TRUE), 
            levels = pretty(zlim, nlevels), nlevels = 20, color.palette = cm.colors, 
            col = color.palette(length(levels) - 1), plot.title, plot.axes, 
            key.title, key.axes, asp = NA, xaxs = "i", yaxs = "i", las = 1, 
            axes = TRUE, frame.plot = axes,mar, ...) 
{
  # modification by Ian Taylor of the filled.contour function
  # to remove the key and facilitate overplotting with contour()
  # further modified by Carey McGilliard and Bridget Ferris
  # to allow multiple plots on one page

  if (missing(z)) {
    if (!missing(x)) {
      if (is.list(x)) {
        z <- x$z
        y <- x$y
        x <- x$x
      }
      else {
        z <- x
        x <- seq.int(0, 1, length.out = nrow(z))
      }
    }
    else stop("no 'z' matrix specified")
  }
  else if (is.list(x)) {
    y <- x$y
    x <- x$x
  }
  if (any(diff(x) <= 0) || any(diff(y) <= 0)) 
    stop("increasing 'x' and 'y' values expected")
 # mar.orig <- (par.orig <- par(c("mar", "las", "mfrow")))$mar
 # on.exit(par(par.orig))
 # w <- (3 + mar.orig[2]) * par("csi") * 2.54
 # par(las = las)
 # mar <- mar.orig
 plot.new()
 # par(mar=mar)
  plot.window(xlim, ylim, "", xaxs = xaxs, yaxs = yaxs, asp = asp)
  if (!is.matrix(z) || nrow(z) <= 1 || ncol(z) <= 1) 
    stop("no proper 'z' matrix specified")
  if (!is.double(z)) 
    storage.mode(z) <- "double"
    # modified from the original code: changed from .Internal(filledcontour(...)) to:
  .filled.contour(as.double(x), as.double(y), z, as.double(levels), 
                          col = col)
  if (missing(plot.axes)) {
    if (axes) {
      title(main = "", xlab = "", ylab = "")
      Axis(x, side = 1)
      Axis(y, side = 2)
    }
  }
  else plot.axes
  if (frame.plot) 
    box()
  if (missing(plot.title)) 
    title(...)
  else plot.title
  invisible()
}
```



```{r profit_spatial, cache=TRUE, include= FALSE}
counties <- readOGR(dsn="shapefiles/county.shp",layer="county")
names(counties)[9] <- "fips" 

profit_county <- tbl(HeatonAgsolver_db, "07_swg_area_counties_tweak") %>%
  as.data.frame() 

 for (i in 1:length(profit_county$fips)) {
   profit_county$fips[i] <- sub("IA", "19", profit_county$fips[i])
    }

counties@data <- left_join(counties@data, profit_county, by = "fips")
counties_df <- fortify(counties)
counties$id <- row.names(counties) 
counties_profit_df <- left_join(counties_df, counties@data)

# add a spatial data frame with the county centroids, to be used to plot the total leaching N per county:
centroids <- getSpPPolygonsLabptSlots(counties) %>%
  as.data.frame() 

centroids$id <- as.character(0:98)

# join the total N leaching data to the centroids 
counties_profit_df <- left_join(counties_profit_df, centroids)

# create a theme without axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )
```



```{r figure_profit, fig.height=4, fig.width=9,  fig.cap="Figure 1: Areas below a mean 2012-2015 profitability of US$ -100 ha^-1^ that are leaching > 50 kg N ha^-1^, per county. Purple shades indicate the relative area, orange circles show the absolute size of the area per county."}
ggplot() +
  geom_polygon(data = counties_profit_df, aes(long, lat, group = group, fill=area_in_swg_perc), color = "white") +
  geom_point(data = counties_profit_df, aes(V1, V2, size = area_in_swg), color = "orange") +
  labs(size = expression(Total~area~losing~">100"~"US$"~ha^{-1}~and~">50"~kg~N~ha^{-1}~(ha))) +
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(Relative~area~losing~">100"~"US$"~ha^{-1}~and~">50"~kg~N~ha^{-1}~("%"))) +
  scale_fill_gradient(low = "white", high="dark blue", guide = "colourbar") 
```

```{r swg_areas_two_thresholds_prepare, cache=TRUE}
# In this chunk, the areas in switchgrass are calculated for all possible combinations of 
# profitability and N loss cut offs.

# swg_areas <- vector(mode="numeric", length=0)
profit_cutoffs_16 <- c(-800, -700, -600, -500, -400, -300, -200, -100, 0, 100, 200, 300, 400, 500, 600, 700)
n_loss_cutoffs_16 <- c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150)

# make a cartesian product of the two vectors first:
# function (from https://www.r-bloggers.com/nested-loops-with-mapply/)
CartProduct = function(vector_1, vector_2)
{
 
  if (length(dim(vector_2)) != 0 )
  {
    warning("New vector has more than one dimension.")
    return (NULL)
  }
 
  if (length(dim(vector_1)) == 0)
  {
    CurrentRows = length(vector_1)
    vector_1 = as.matrix(vector_1, nrow = CurrentRows, ncol = 1)
  } else {
    CurrentRows = nrow(vector_1)
  }
 
  var1 = replicate(length(vector_2), vector_1, simplify=F)
  var1 = do.call("rbind", var1)
 
  var2 = rep(vector_2, CurrentRows)
  var2 = matrix(var2[order(var2)], nrow = length(var2), ncol = 1)
 
  CartProduct = cbind(var1, var2)
  return (CartProduct)
}

comb <- CartProduct(profit_cutoffs_16, n_loss_cutoffs_16)

# write a function that uses each of the combinations, using the matrix returned from the function CartProduct:
area_select <- function(cutoff_1, cutoff_2)
{ query <- paste("SELECT SUM(clumuha) FROM \"05_dndc_clumu_cgsb_swg\" WHERE mean_profit_ha < ", as.character(cutoff_1), "AND ave_no3_leach_ha_cgsb > ", as.character(cutoff_2))
    sum_area <- tbl(HeatonAgsolver_db, sql(query)) %>%
      as.data.frame()
    sum_area <- sum_area[1,1]
    return(sum_area)
}
```

```{r swg_areas_two_thresholds, cache=TRUE}
# apply the function to each value in the matrix
swg_areas <- mapply(area_select, comb[,1], comb[,2])


# combine the cut off values with the area in switchgrass in a data frame
swg_areas <- data.frame(profit_cut = comb[,1],
                        no3_leach_cut = comb[,2],
                        swg_areas) 

```

```{r no3_leach_function, cache=TRUE}

# write a function that uses each of the cutt off combinations and calculates the total NO3 leaching in this scenario:
no3_leach_select <- function(cutoff_1, cutoff_2)
{query <- paste("SELECT (sum((CASE WHEN mean_profit_ha <" , as.character(cutoff_1), " AND ave_no3_leach_ha_cgsb > ",  as.character(cutoff_2), " THEN ave_no3_leach_ha_swg_10000  ELSE ave_no3_leach_ha_cgsb END) * clumuha)) / 1000 FROM \"05_dndc_clumu_cgsb_swg\" ")
    sum_no3_leach <- tbl(HeatonAgsolver_db, sql(query)) %>%
      as.data.frame()
    sum_no3_leach <- sum_no3_leach[1,1]
    return(sum_no3_leach)
}

# this is the SQL query that selects for each polygon the value of cgsb or swg, depending on its cgsb NO3 leaching and profit value, 
# and sums all values up 
# to get a Iowa wide NO3 leaching value (absolute), that can then substracted from the baseline NO3 leaching.

```
```{r no3_leach_two_thresholds, cache=TRUE}

# apply the function to each value in the matrix
tot_no3_leach <- mapply(no3_leach_select, comb[,1], comb[,2])

# combine the cut off values with the area in switchgrass in a data frame
swg_no3_leach <- data.frame(profit_cut = comb[,1],
                        no3_leach_cut = comb[,2],
                        tot_no3_leach) 


```


```{r area_surfaceplot, cache = TRUE, fig.height=9, fig.width=10,  fig.cap="Figure 2: State-wide aggregated model outputs for (a) area converted to switchgrass, (b) biomass production, assuming a 10,000 kg ha^-1^ yield, and (c) reduction N loss through nitrate leaching as a function of profitability and nitrate leaching driven management decisions. Each corner in the surface grid represents an output value under given combination of nitrate leaching threshold (the value in kg N ha^-1^ above which the area would be converted to switchgrass) and profitability threshold (the value in US$ ha^-1^ below which the area would be converted to switchgrass)."}

x <- profit_cutoffs_16
y <- n_loss_cutoffs_16

# create a matrix out of the z values (in Mha):
swg_matrix <- matrix(swg_areas[,3] * 1e-6
, nrow=16, ncol=16)
#calculate % of total cropland (9374363 ha)
tot_area <- 9.374363
swg_matrix_perc <- (swg_matrix/tot_area)*100
z <- swg_matrix_perc

# create figure with three plots 
plot.new()

#I am organizing where the plots appear on the page using the "plt" argument in "par()"

# top left plot:
par(new = "TRUE",              
    plt = c(0.05,0.45,0.55,0.9),   # using plt instead of mfcol (compare
                                  # coordinates in other plots)
    las = 1,                      # orientation of axis labels
    cex.axis = 1,                 # size of axis annotation
    tck = 0.02 )                  # major tick size and direction, < 0 means outside

persp3D(x, y, z, colvar = z, phi = 10, theta = -150, col = brewer.pal(8,"YlGn"), 
        border = 17, facets = TRUE, colkey = TRUE, resfac = 1, clab = "Switchgrass\narea (%)", 
        lighting = list(alpha = 1),
        xlab = "Profit threshold", ylab = "N loss threshold", zlab = "Swg area (%)",
ticktype = "detailed")


text3D(x=900,y=150, z=109, "(a)", cex = 1.5, font = 2, phi = 10, theta = -150, add= TRUE)


#calculate total biomass for Iowa for each scenario (10,000 kg/ha) in million Mg
swg_matrix_bm <- (swg_matrix * 10)
z_bm <- swg_matrix_bm

#Top right plot:
par(new = "TRUE",
    plt = c(0.55,0.95,0.55,0.9),  # defining window for second plot
    las = 1,
    cex.axis = 1)

persp3D(x, y, z_bm, colvar = z_bm, phi = 10, theta = -150, col = brewer.pal(8,"YlGn"), 
        border = 17, facets = TRUE, colkey = TRUE, resfac = 1, clab = "Biomass\nproduction \n(million metric tons)", 
        lighting = list(alpha = 1),
        xlab = "Profit threshold", ylab = "N loss threshold", zlab = "Biomass (million tons)",
        ticktype = "detailed")

text3D(x=900,y=150, z=102, "(b)", cex = 1.5, font = 2, phi = 10, theta = -150, add= TRUE)




# create matrix for z values from the data frame swg_no3_leach (N leaching loss in kg)
swg_matrix_nl <- matrix(swg_no3_leach[,3], nrow = 16, ncol = 16)
#calculate N loss reduction compared to baseline
no3_leach_tot <-tbl(HeatonAgsolver_db, "05_dndc_no3_leach_sums_iowa_scenarios")  %>%
  as.data.frame()
cgsb_no3_leach <- no3_leach_tot[1,1]
swg_matrix_nlr <- (cgsb_no3_leach - swg_matrix_nl) *100 / cgsb_no3_leach
z_nlr <- swg_matrix_nlr

#Bottom left plot:
par(new = "TRUE",
    plt = c(0.05,0.45,0.05,0.4),
    las = 1,
    cex.axis = 1)

persp3D(x, y, z_nlr, colvar = z_nlr, phi = 10, theta = -150, col = brewer.pal(8,"YlGn"), 
        border = 17, facets = TRUE, colkey = TRUE, resfac = 1, clab = "N loss \nreduction (%)", 
        lighting = list(alpha = 1),
        xlab = "Profit threshold", ylab = "N loss threshold", zlab = "N loss reduction (%)",
ticktype = "detailed")

text3D(x=900,y=150, z=85, "(c)", cex = 1.5, font = 2, phi = 10, theta = -150, add= TRUE)





```

```{r area_contourplot, cache = TRUE, fig.height=7, fig.width=8, fig.cap="Figure 3: Contour plots (vertical projections of Figure 2) showing (a) the resulting areas converted to switchgrass (in % of row cropland), (b) total switchgrass biomass production (in million metric tons), and (c) total nitrate leaching reduction (in % of the baseline) as a function of nitrate leaching and profitability threshold values of the baseline scenario. In each plot, isolines with numbers 10-70 show the threshold combinations with similar resulting values. The grey dashed lines show an example how to read the plot: If cropland losing > 50 kg N ha^-1^ and losing > US$ 100 ha^-1^ was considered for conversion to switchgrass (tweak scenario), a little over 10% of cropland currently in rowcrop would be in switchgrass (a), approx. 10 million metric tons of biomass would be produced (b), and the N loss through leaching would be reduced by almost 20 % (c). the solid lines show the outcomes for the nutrient reduction scenario."}


# multiple panel plot with the modified function filled.contour3()

plot.new()

#I am organizing where the plots appear on the page using the "plt" argument in "par()"

# top left plot:
par(new = "TRUE",              
    plt = c(0.1,0.45,0.60,0.95),   # using plt instead of mfcol (compare
                                  # coordinates in other plots)
    las = 1,                      # orientation of axis labels
    cex.axis = 1,                 # size of axis annotation
    tck = -0.02 )                  # major tick size and direction, < 0 means outside

mylevels <- c(seq(0,70,10),100)
filled.contour3(x, y, z, 
                col = brewer.pal(length(mylevels),"YlGn"), 
                levels=mylevels, 
                xlab="", 
                ylab=expression(N~loss~threshold~(kg~ha^{-1})),
                xlim = c(-400,700), ylim = c(0,120))
contour(x, y, z, 
        levels=mylevels, 
        add = TRUE, 
        xlim = c(-400,700), ylim = c(0,120))
par(xpd = FALSE) # restricts the lines to the plot region
abline(h = 50, col = "dark gray", lty = 2)
abline(v = -100, col = "dark gray", lty = 2)
abline(h = 20, col = "dark gray", lty = 1)
abline(v = 0, col = "dark gray", lty = 1)

# An annotation inside first plot
#The xpd=NA allows for writing outside the plot limits, but still using the the x and y axes to place the text
par(xpd = NA)
text(x=-300,y=110,"(a)",cex = 1.5,font = 2)


#Top right plot:
par(new = "TRUE",
    plt = c(0.55,0.9,0.60,0.95),  # defining window for second plot
    las = 1,
    cex.axis = 1)

filled.contour3(x, y, z_bm, 
                col = brewer.pal(length(mylevels),"YlGn"), 
                levels=mylevels, 
                xlab=expression(Profitability~threshold~("US$"~ha^{-1})), 
                ylab="",
                xlim = c(-400,700), ylim = c(0,120))
contour(x, y, z_bm, 
        levels=mylevels, 
        add = TRUE, 
        xlim = c(-400,700), ylim = c(0,120))
par(xpd = FALSE) # restricts the lines to the plot region
abline(h = 50, col = "dark gray", lty = 2)
abline(v = -100, col = "dark gray", lty = 2)
abline(h = 20, col = "dark gray", lty = 1)
abline(v = 0, col = "dark gray", lty = 1)

par(xpd = NA)
text(x=-300,y=110,"(b)",cex = 1.5,font = 2)


#Bottom left plot:
par(new = "TRUE",
    plt = c(0.1,0.45,0.15,0.5),
    las = 1,
    cex.axis = 1)

filled.contour3(x, y, z_nlr, 
                col = brewer.pal(length(mylevels),"YlGn"), 
                levels=mylevels, 
                xlab=expression(Profitability~threshold~("US$"~ha^{-1})), 
                ylab=expression(N~loss~threshold~(kg~ha^{-1})), 
                xlim = c(-400,700), ylim = c(0,120))
contour(x, y, z_nlr, 
        levels=mylevels, 
        add = TRUE, 
        xlim = c(-400,700), ylim = c(0,120))
par(xpd = FALSE) # restricts the lines to the plot region
abline(h = 50, col = "dark gray", lty = 2)
abline(v = -100, col = "dark gray", lty = 2)
abline(h = 20, col = "dark gray", lty = 1)
abline(v = 0, col = "dark gray", lty = 1)

par(xpd = NA)
text(x=-300,y=110,"(c)",cex = 1.5,font = 2)


```

```{r counties_attributes, cache=TRUE}
 counties_no3_leach = tbl(HeatonAgsolver_db, "07_dndc_counties") %>%
   as.data.frame()
 # change the data in the column named "FIPS", so that is contains the county identifiers in the same format as the spatial file:
 for (i in 1:length(counties_no3_leach$fips)) {
   counties_no3_leach$fips[i] <- sub("IA", "19", counties_no3_leach$fips[i])
    }
 # make it a factor
  counties_no3_leach$fips <- as.factor(counties_no3_leach$fips)
```

```{r counties_spatial, cache=TRUE, include=FALSE, cache=TRUE}
counties <- readOGR(dsn="shapefiles/county.shp",layer="county")
names(counties)[9] <- "fips" # change to lower case so that we can join on this field later on

# join the attribute data with the spatial data, using a left join and the identifier "fips":
counties@data <- left_join(counties@data, counties_no3_leach, by = "fips")
counties_df <- fortify(counties)
counties$id <- row.names(counties) # assign a column with the variable ID that is equal to the row names (=numbers), starting with 0.

```

```{r counties_centroids, warning=FALSE, message=FALSE, cache=TRUE} 
# add a spatial data frame with the county centroids, to be used to plot the total leaching N per county:
centroids <- getSpPPolygonsLabptSlots(counties) %>%
  as.data.frame() 
centroids$id <- as.character(0:98)
```

```{r data_n_loss_maps, cache=TRUE}
# set up the data frame 
counties_swg <- data.frame(id = rep(counties$id,3),
                         scenario = c(rep("swg_7500", length(counties$id)), rep("swg_10000", length(counties$id)), 
                                                                               rep("swg_12500", length(counties))),
                         tot_no3_leach = c(counties$tot_no3_leach_7500, counties$tot_no3_leach_10000,
                                           counties$tot_no3_leach_12500),
                         ave_no3_leach = c(counties$ave_no3_leach_7500, counties$ave_no3_leach_10000, counties$ave_no3_leach_12500),
                         no3_leach_red = c(counties$no3_leach_change_7500, counties$no3_leach_change_10000,
                                       counties$no3_leach_change_12500))

counties_swg$scenario <- factor(counties_swg$scenario, levels = c("swg_7500", "swg_10000", "swg_12500"))
# join data frame with the data frame extracted out of the spatial data frame
counties_swg_df <- left_join(counties_df, counties_swg) %>%
  left_join(centroids)
```

```{r table_no3_leach_reduction, cache=TRUE}
no3_leach_tot <-tbl(HeatonAgsolver_db, "05_dndc_no3_leach_sums_iowa_scenarios")  %>%
  as.data.frame()
no3_leach_tot <- no3_leach_tot[1:4]
no3_leach_tot <-   data.frame(value = "Total leaching (Mg N)", 
                           no3_leach_tot)
names(no3_leach_tot) <- c(" ", "Status Quo", "7500", "10000", "12500")

no3_leach_red <- data.frame(value = "Reduction (%)", 
                         status_quo = 0,
                        "7500" = (no3_leach_tot[1,2]-no3_leach_tot[1,3])*100/no3_leach_tot[1,2],
                        "10000" = (no3_leach_tot[1,2]-no3_leach_tot[1,4])*100/no3_leach_tot[1,2],
                        "12500" = (no3_leach_tot[1,2]-no3_leach_tot[1,5])*100/no3_leach_tot[1,2])
names(no3_leach_red) <- c(" ", "Status Quo", "7500", "10000", "12500")

no3_leach_table <- rbind(no3_leach_tot[1,],no3_leach_red[1,])

kable(no3_leach_table, digits = 2, col.names=c(" ", "Status Quo","7500", "10000", "12500"), caption="Table 1: Annual N loss through leaching (Mg N) and nitrate leaching reduction (%) in Iowa modelled with DNDC under different switchgrass yield scenarios (7,500, 10,000, and 12,500 kg ha^-1^)." )

```

```{r plot_n_loss_map, cache = TRUE, fig.width=8, fig.height=4, fig.cap = "Figure 4: Mean (green shades) and total (yellow circles) annual nitrate leaching per county, averaged over 2012-2015 for the status quo cropland."}

counties_swg_med_df <- filter(counties_swg_df, scenario == "swg_10000")

ggplot(counties_swg_med_df) +
  geom_polygon(aes(long, lat, group = group, fill=ave_no3_leach), color = "white") +
  scale_fill_gradient(low = "white", high="#009E73", guide = "colourbar") +  
  labs(fill = expression(atop("Average nitrate", paste("leaching (kg N ha"^{-1},")")))) +  
  geom_point(aes(V1, V2, size = tot_no3_leach), color = "yellow") +
  labs(size = expression(Total~leached~nitrate~(Mg~N))) +
  scale_size(range = c(0, 6)) +
  coord_equal() +
  ditch_the_axes 
```

```{r plot_n_loss_reduction_map, cache = TRUE, fig.width=8, fig.height=4, fig.cap= "Figure 5: Relative nitrogen leaching reduction per county when switchgrass is grown on areas that loose > US$ 100 ha^-1^ and > 50 kg N ha^-1^, assuming medium (10,000 kg ha^-1^) switchgrass yield, as compared to the status quo in Figure 4."}
ggplot(counties_swg_med_df) +
  geom_polygon(aes(long, lat, group = group, fill=no3_leach_red), color = "white") +
  scale_fill_gradient(low = "white", high="#0072B2", guide = "colourbar") +  
  coord_equal() +
  ditch_the_axes +
  labs(fill = expression(Nitrate~leaching~reduction~("%"))) 
```


```{r subfield_n_red_iowa_zoom, cache = TRUE, fig.height=5, fig.width=5, fig.cap="Figure 6: Subfield-scale relative nitrate leaching reduction in Iowa ((a), (b)) and in the northern part of the North Raccoon River Watershed, spanning across Buena Vista, Pocahontas, Sac, and Calhoun County ((c), (d)). (a) and (c) show the \"tweak scenario\" where switchgrass is grown on areas that loose > US$ 100 ha^-1^ and > 50 kg N ha^-1^. (b) and (d) show the \"nutrient reduction scenario\" where switchgrass is grown on unprofitable areas that leach > 20 kg N ha^-1^."}

include_graphics("imported_images/N_leaching_subfield_RRWS.tif", dpi = 300)
```

## Discussion

### Comparison of results to measured values
In comparison to a maize-maize-soybean rotation, miscanthus, switchgrass, and prairie reduced nitrate leaching at 50 cm soil depth by more than 90% in the 4th year [@smith2013reduced].  @daigh2015subsurface showed that fertilized, reconstructed prairie lost 89-99% less nitrate N in tile drainage water compared to row crops. @mcisaac2010miscanthus showed a 97% reduction in nitrate leaching at 50 cm soil depth.

- we are not considering tile drainage, because the model predicts leaching at 50 cm soil depth (assumed to be below the rooting zone). It can be assumed that further nitrate is being removed from the soil below 50 cm, especially from deep routing switchgrass plants.

- atmospheric deposition of N is not considered by the model (? - check), but it is factored out when calculating N loss changes

- manure application is not considered in the model due to missing spatial data on manure application and type

- our results do not take into account denitrification in surface waters, so the sum is not equal to the DIN export to the Gulf of Mexico [@alexander2008differences]

- however, by estimating DIN loss on a field basis, we are accounting for the amount of dissolved inorganic nitrogen that is lost for the farmer

- *compare with North Racoon River watershed data in @schilling2000relationship: They showed N loads at the outlet of the watershed*

- *can I compare our results with N yields of watersheds, such as reported by @alexander2008differences? What is the total flux per state in table S4 (no unit, seems to be % of total flux into the Gulf, so how much is that???)*

- integration of switchgrass on 10% of cropland would have a stronger effect on N loss reduction compared to other conservation practices, such as crop rotation or conservation tillage [@wu2004decisions]

To reach the Iowa Nutrient Reduction Strategy's goal of 45% N leaching reduction, 40-50% of Iowa cropland would have to be converted to switchgrass. Our resuls indicate the need to combine multiple conservation practices and to adapt them to the biophysical and socio-economic conditions of each given region.

- the variability accross the state comes from the variability in candidate areas for conversion, and from the effectivenitrate leaching reduction of each of these areas

- outlook: @qiu2015landscape showed that landscape composition is affecting water quality related ecosystem services to varying extent. Due to the large spatial area covered, our underlying data includes a realistic but approximated reflection of reality. However, applied to smaller management units and fed with more detailed and accurate data, our spatial results could be used as indicators to identify nonlinear (disproportionate) responses of N loss reduction to conversion of land to perennial grasses.
(--> find "sweet spots" where large improvements can be obtained by conversion of small land areas)


- annual variability in N leaching due to weather (same with flux values, see @sprague2011nitrate)
## Conclusions


- This study disaggregates previously published approaches of N balance estimated based on county data (e.g Libra et al 2004) down to a finer resolution and takes into account the biogeochemical processes driving DIN loss in agroecosystems

## Supplementary Information

```{r weather_data, cache = TRUE}

# make a list of files and their paths that I want to load, using regular expressions to filter according to a pattern
wth <- list.files(path = "./weather", pattern = "IA[[:digit:]][[:digit:]][[:digit:]]_201[2345].wth", full.names = TRUE)

# make a vector from wth that only includes the fips and year
fips <- list.files(path = "./weather", pattern = "IA[[:digit:]][[:digit:]][[:digit:]]_201[2345].wth")
# split the file name to get the fips and create a matrix with two columns (there must be a way to extract only the fips but I don't know it)
fips <- unlist(strsplit(fips, "_")) %>%
matrix(ncol = 2, byrow = TRUE)
# create vector to use for the final data frame
fips <- fips[,1]

# create a year vector to use for the final data frame
year <- rep(c("2012", "2013", "2014", "2015"), length(fips)/4)

# add a header for the files   
header <- c("doy", "temp_max", "temp_min", "precip" )

# read in the files as a list of data frames and then combine them into one data frame
weather <- lapply(wth, read.table, skip = 1, col.names = header) %>%
  bind_rows() %>%
  mutate(fips = rep(fips, each = 365 )) %>%
  mutate(year = rep(year, each = 365))


# using dplyr, I can now aggregate the data to county, year, etc.
weather %>%
  group_by(fips, year) %>%
  summarise(sum(precip)) %>%
  rename(precip = `sum(precip)`) -> annual_precip

# extracting the relevant data from the postgreSQL database
query_annual <- "SELECT fips, ave_no3_leach12_cgsb, ave_no3_leach13_cgsb, ave_no3_leach14_cgsb, ave_no3_leach15_cgsb FROM \"07_dndc_counties\" "
  
annual_leach <- tbl(HeatonAgsolver_db, sql(query_annual)) %>%
      as.data.frame() %>%
      rename("2012" = ave_no3_leach12_cgsb, "2013" = ave_no3_leach13_cgsb, "2014" = ave_no3_leach14_cgsb, "2015" = ave_no3_leach15_cgsb) %>%
      gather(year, avg_leach, `2012`, `2013`, `2014`, `2015`)  %>% # gather the table to get the variable "no3 leaching" into one column (use back ticks for years)
      left_join(annual_precip, by = c("fips", "year")) -> leach_precip

leach_precip$year <- as.factor(leach_precip$year)
```

```{r plot_leach_precip}

ggplot(leach_precip) +
  geom_point(aes(x = precip, y = avg_leach, col = year)) +
  theme_b_border() +
  labs(x=expression(Annual~precipitation~(mm)),
       y=expression(Annual~nitrate~leaching~(kg~N~ha^{-1}))) +
  scale_x_continuous() +
  scale_y_continuous() +
    scale_color_discrete(name="Year")

```



```{r profit_distribution, fig.width=4.5, fig.height=3, fig.cap="Figure S1: Distribution of mean profitability 2012-2015. The vertical line marks the US$ -100  ha^-1^  cut off value used for the switchgrass integration analysis in Figure S5."}
profit = tbl(HeatonAgsolver_db, "profit_mean_2012_2015_aggregated") %>%
  as.data.frame()
profit$area_ha <- profit$area_ha * 1e-6  # convert to Mha

ggplot(profit, aes(profit_ha_rounded, weight = area_ha)) + 
  geom_histogram(binwidth=20, alpha = .5, position="identity") +
  theme_b_border() +
  labs(x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1})),
       y=expression(Area~(Mha))) +
  scale_x_continuous() +
  scale_y_continuous() +
  geom_vline(aes(xintercept= -100, linetype= "dashed"),  show.legend = FALSE) 

```

```{sql, connection=con, cache=TRUE, include=FALSE}
SELECT setseed(0.5)
```

```{sql, connection=con, cache=TRUE}
CREATE TABLE "r_dndc_clumu_cgsb_swg_sample" AS SELECT mean_profit_ha, ave_no3_leach_ha_cgsb FROM "05_dndc_clumu_cgsb_swg" ORDER BY random() LIMIT 10000
```
```{r data_sample_n_loss}
sample_leach <- tbl(HeatonAgsolver_db, "r_dndc_clumu_cgsb_swg_sample") %>%
  as.data.frame(sample)
```

```{r scatterplot_no3_leach, fig.width=4.5, fig.height=3, fig.cap="Figure S2: Scatterplot of a subsample (n=10,000) of profitability and nitrate leaching on subfield areas of the baseline scenario (mean values for maize and soybean fields 2012-2015) in Iowa. The red lines indicate the cut off values for the minimum (tweak) scenario (dashed lines, US$ -100 ha^-1^ for profitability and 50 kg N ha^-1^ for nitrate leaching) and for the nutrient reducton scenario (solid lines,  US$ 0 ha^-1^ for profitability and 20 kg N ha^-1^ for nitrate leaching)."}
ggplot() +
  geom_point(data = sample_leach, aes(x = mean_profit_ha, y = ave_no3_leach_ha_cgsb), alpha = .2, stroke=0) +
  theme_b_border() +
  labs(y=expression(Mean~NO[3]^{"-"}~leaching~2012-2015~(kg~N~ha^{-1})),
       x=expression(Mean~profitability~2012-2015~("US$"~ha^{-1}))) +
  geom_vline(xintercept= -100, linetype= "dashed", color = "red",  show.legend = FALSE) +
  geom_hline(yintercept= 50, linetype= "dashed", color = "red",  show.legend = FALSE) +
    geom_vline(xintercept= 0, linetype= "solid", color = "red",  show.legend = FALSE) +
  geom_hline(yintercept= 20, linetype= "solid", color = "red",  show.legend = FALSE)
  
```

```{r subfield_n_loss_maps, fig.height=5.981, fig.width=4.592, fig.cap="Figure S3: (a) Nitrate leaching and (b) ammonium volatilization modelled with DNDC on a subfield resolution for Iowa. Average values for 2012-2015 are shown."}

include_graphics("imported_images/Nitrogen_subfield_2012_2015_raster.tif", dpi = 300)

```


```{r fetch_distr, cache=TRUE}

cgsb_n_loss <-tbl(HeatonAgsolver_db, "06_cgsb_ave_n_loss_rounded_aggr")  %>%
  as.data.frame()
cgsb_no3_leach <-tbl(HeatonAgsolver_db, "06_cgsb_ave_no3_leach_rounded_aggr")  %>%
  as.data.frame()
cgsb_nh3_vol <-tbl(HeatonAgsolver_db, "06_cgsb_ave_nh3_vol_rounded_aggr")  %>%
  as.data.frame()
swg_low <-tbl(HeatonAgsolver_db, "06_swg_7500_ave_n_loss_rounded_aggr")  %>%
  as.data.frame()
swg_med <-tbl(HeatonAgsolver_db, "06_swg_10000_ave_n_loss_rounded_aggr")  %>%
  as.data.frame()
swg_hi <-tbl(HeatonAgsolver_db, "06_swg_12500_ave_n_loss_rounded_aggr")  %>%
  as.data.frame()
```
```{r data_list}
data_list <- list(cgsb_n_loss, cgsb_no3_leach, cgsb_nh3_vol, swg_low, swg_med, swg_hi)

scenario <- c(rep("cgsb_n_loss", sum(sapply(data_list[1],nrow))), 
              rep("cgsb_no3_leach", sum(sapply(data_list[2],nrow))),
              rep("cgsb_nh3_vol", sum(sapply(data_list[3],nrow))), 
              rep("swg_low", sum(sapply(data_list[4],nrow))), 
              rep("swg_med", sum(sapply(data_list[5],nrow))), 
              rep("swg_hi", sum(sapply(data_list[6],nrow)))) %>%
factor(levels= c('cgsb_n_loss', 'cgsb_no3_leach', 'cgsb_nh3_vol', 'swg_low', 'swg_med', 'swg_hi')) # make it a factor so that it is plotted in the right order in the facets


leaching <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   leaching <- append(leaching, data_list[[i]][,1])
area <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   area <- append(area, data_list[[i]][,2])
   
leach <- data.frame(scenario, leaching, area = area*1e-6) # this factor to show Mha

# calculate weighted medians and means

# median_cgsb <- weightedMedian(cgsb_leach[-1,1], w=cgsb_leach[-1,2])
# median_swg_low <- weightedMedian(swg_low[,1], w=swg_low[,2])
# median_swg_med <- weightedMedian(swg_med[,1], w=swg_med[,2])
# median_swg_hi <- weightedMedian(swg_hi[,1], w=swg_hi[,2])
# medians <- c(median_cgsb, median_swg_low, median_swg_med, median_swg_hi)
# medians_df <- data.frame(crop = levels(leach$scenario),
#                          median = medians)
# 
# mean_cgsb <- weightedMean(cgsb_leach[-1,1], w=cgsb_leach[-1,2])
# mean_swg_low <- weightedMean(swg_low[,1], w=swg_low[,2])
# mean_swg_med <- weightedMean(swg_med[,1], w=swg_med[,2])
# mean_swg_hi <- weightedMean(swg_hi[,1], w=swg_hi[,2])
# means <-c(mean_cgsb, mean_swg_low, mean_swg_med, mean_swg_hi)


```

```{r histograms, cache=TRUE, fig.height=4, fig.width=7, fig.cap="Figure S4: State-wide distributions of N loss from maize and soybean fields (upper panels: total, by nitrate leaching only, by ammonia volatilization only) and mean nitrate leaching loss if all cropland was converted to switchgrass (lower panels, assuming low, medium and high switchgrass yields). Values for switchgrass were averaged over the mananagement horizon of 10 years , including two establishment years (2001 - 2015."}
scenarios <- c('cgsb_n_loss'= "N loss (corn/soybeans)",'cgsb_no3_leach'= "NO3 leach. (corn/soybeans)", 'cgsb_nh3_vol'= "NH3 vol. (corn/soybeans)",'swg_low' = "Low yielding switchgrass", 'swg_med' = "Medium yielding switchgrass", 'swg_hi' = "High yielding switchgrass")

ggplot(leach, aes(leaching, weight = area)) + 
  geom_histogram(binwidth=2, alpha = .5, position="identity") +
  theme_b_border() +
  labs(x=expression(Mean~nitrogen~loss~(kg~N~ha^{-1})),
       y=expression(Area~(Mha))) +
  scale_x_continuous(lim = c(-1,150)) +
  scale_y_continuous() +
 # geom_vline(data = medians_df, aes(xintercept=median), color = "black") +
  theme(legend.position="bottom") +
  facet_wrap( ~ scenario, ncol=3, labeller = as_labeller(scenarios)) 
```

```{r swg_areas, cache=TRUE}
swg_areas_11 <- vector(mode="numeric", length=0)
profit_cutoffs <- c(-500, -400, -300, -200, -150, -100, -50, 0, 50, 100, 150)
 for (i in 1:length(profit_cutoffs)) {
    query <- paste("SELECT SUM(clumuha) FROM \"05_dndc_clumu_cgsb_swg\" WHERE mean_profit_ha < ", as.character(profit_cutoffs[i]), "AND ave_no3_leach_ha_cgsb > 50")
    sum_area <- tbl(HeatonAgsolver_db, sql(query)) %>%
      as.data.frame()
    sum_area <- sum_area[1,1]
    swg_areas_11 <- append(swg_areas_11, sum_area) 
  }
```

```{r query_leach_sums, cache=TRUE}

# query sum of leaching from the baseline (corn/beans) scenario
query_cgsb <- "SELECT(sum(ave_no3_leach_ha_cgsb * clumuha)) / 1000 FROM \"05_dndc_clumu_cgsb_swg\" "
leach_cgsb <- tbl(HeatonAgsolver_db, sql(query_cgsb)) %>%
      as.data.frame()
leach_cgsb <- leach_cgsb[1,1] 

# query sums of leaching from the low yield swg integration scenario, with constant leaching cut off and varying profit cut off
leach_sums_lo <- vector(mode="numeric", length = 0)
for (i in 1:length(profit_cutoffs)) {
  query <- paste("SELECT (sum((CASE WHEN mean_profit_ha <", as.character(profit_cutoffs[i]),"AND ave_no3_leach_ha_cgsb > 50 THEN ave_no3_leach_ha_swg_7500 ELSE ave_no3_leach_ha_cgsb END) * clumuha)) / 1000 FROM \"05_dndc_clumu_cgsb_swg\" ")
  leach_sum <- tbl(HeatonAgsolver_db, sql(query)) %>%
      as.data.frame()
  leach_sum <- leach_sum[1,1]
  leach_sums_lo <- append(leach_sums_lo, leach_sum)
}


# query sums of leaching from the med yield swg integration scenario, with constant leaching cut off and varying profit cut off
leach_sums_med <- vector(mode="numeric", length = 0)
for (i in 1:length(profit_cutoffs)) {
  query <- paste("SELECT (sum((CASE WHEN mean_profit_ha <", as.character(profit_cutoffs[i]),"AND ave_no3_leach_ha_cgsb > 50 THEN ave_no3_leach_ha_swg_10000 ELSE ave_no3_leach_ha_cgsb END) * clumuha)) / 1000 FROM \"05_dndc_clumu_cgsb_swg\" ")
  leach_sum <- tbl(HeatonAgsolver_db, sql(query)) %>%
      as.data.frame()
  leach_sum <- leach_sum[1,1]
  leach_sums_med <- append(leach_sums_med, leach_sum)
}


# query sums of leaching from the high yield swg integration scenario, with constant leaching cut off and varying profit cut off
leach_sums_hi <- vector(mode="numeric", length = 0)
for (i in 1:length(profit_cutoffs)) {
  query <- paste("SELECT (sum((CASE WHEN mean_profit_ha <", as.character(profit_cutoffs[i]),"AND ave_no3_leach_ha_cgsb > 50 THEN ave_no3_leach_ha_swg_12500 ELSE ave_no3_leach_ha_cgsb END) * clumuha)) / 1000 FROM \"05_dndc_clumu_cgsb_swg\" ")
  leach_sum <- tbl(HeatonAgsolver_db, sql(query)) %>%
      as.data.frame()
  leach_sum <- leach_sum[1,1]
  leach_sums_hi <- append(leach_sums_hi, leach_sum)
}

```


```{r data_dispr_benefits}
# organize the data in a data frame to be used in ggplot
swg_yield <- as.factor(c("baseline",rep("7500",11), rep("10000",11),rep("12500",11)))
profit_cutoff <- c(NA, rep(profit_cutoffs,3))
swg_area <- c(0,rep(swg_areas_11,3))
swg_area_percent <- (swg_area/9374363)*100
no3_leach <- c(leach_cgsb, leach_sums_lo, leach_sums_med, leach_sums_hi)
leach_red <- abs((no3_leach - no3_leach[1])/no3_leach[1])*100

relation <- data.frame(swg_yield, profit_cutoff, swg_area_percent, leach_red)
```


```{r relations, cache=TRUE, fig.show="hold", fig.width=8, fig.height=3,  fig.cap="Figure S4: State-wide nitrate leaching reduction as a function of area in switchgrass (a) and of the profitability threshold (b). The threshold for nitrate leaching is held constant at 50 kg N ha^-1^ (only areas losing > 50 kg N ha^-1^ through leaching are considered for conversion to switchgrass)."}

plot1 <- ggplot() +
  geom_point(data = relation, aes(x = swg_area_percent, y = leach_red, color=swg_yield)) +
  theme_b_border() +
  labs(x=expression(Area~"in"~switchgrass~("%")),
       y=expression(Nitrate~leaching~reduction~("%"~N))) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete(name="Switchgrass\nYield (kg/ha)",  guide = FALSE) +
  annotate("text", x = 0, y = 40, label = "(a)", size = 5)

plot2 <- ggplot() +
  geom_point(data = relation, aes(x = profit_cutoff, y = leach_red, color=swg_yield)) +
  theme_b_border() +
  labs(x=expression(Profitability~threshold~("US$"~ha^{-1})),
       y=expression(Nitrate~leaching~reduction~("%"~N))) +
  scale_x_continuous(breaks=seq(from=-500, to=200, by= 100)) +
  scale_y_continuous() +
  scale_color_discrete(name="Switchgrass\nYield (kg/ha)") +
  theme(legend.position=c(0.25,0.65))+
  annotate("text", x = -500, y = 40, label = "(b)", size = 5)

grid.arrange(plot1, plot2, ncol = 2)
```
```{r shut, include=FALSE}
 rm(list='HeatonAgsolver_db'); gc() 
```